---
title: "A Gallery of Figure Types"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{A Gallery of Figure Types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7.2,
  fig.height = 4.5,
  dpi        = 192,
  out.width  = "100%",
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while a theme set is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

```{r setup}
library(ggplot2)
library(themeset)

# The packages of the figures are suggested, not required: without one of them
# its chunk is shown but not run.
has_network <- requireNamespace("igraph", quietly = TRUE) && requireNamespace("ggraph", quietly = TRUE)
has_circos <- requireNamespace("circlize", quietly = TRUE)
has_patchwork <- requireNamespace("patchwork", quietly = TRUE)
has_composite <- has_patchwork && has_circos &&
  requireNamespace("ComplexHeatmap", quietly = TRUE) && requireNamespace("ggridges", quietly = TRUE)
```

## Overview

Not every figure is a plot of one variable against another. This article draws
five kinds that papers and reports use often, and styles each of them with
`themeset`:

- a **circular network**, with [ggraph](https://ggraph.data-imaginist.com/)
- a **circos plot** of genomic tracks, with
  [circlize](https://jokergoo.github.io/circlize_book/book/)
- a **four-panel figure** for print, with
  [patchwork](https://patchwork.data-imaginist.com/)
- a **grouped dot plot**, with ggplot2
- a **composite figure** of nine panels, with heat maps and an oncoprint from
  [ComplexHeatmap](https://jokergoo.github.io/ComplexHeatmap-reference/book/)
  next to ggplot2 panels

All the data are simulated, and the numbers in the figures mean nothing. The
packages for the figures are only suggested by `themeset`; if one is not
installed, the code of its figure is shown but not run.

What `themeset` brings is the same in each: the palettes, by name with
`scale_color_set()` or as colors with `journal_colors()`, the theme of a journal
with `journal_theme()`, and shades of one color for ordered groups with
`journal_ramp()`.

---

## 1. A Circular Network

A network of 150 nodes in six groups, drawn on a circle. The nodes of a group
sit side by side, which makes the blocks of the circle. The arcs inside the
circle are the edges, and most of them join two nodes of one group. The size of
a node is its degree, the number of edges that end in it.

```{r network, eval = has_network, fig.width = 7, fig.height = 5.6, fig.cap = "A circular network of 150 nodes in six groups, with arcs for the edges.", fig.alt = "Dots on a circle, in six colored blocks with the name of each block outside the circle. Arcs inside the circle join dots, mostly within a block. The dots differ in size."}
library(ggraph)

set.seed(42)
steps <- c("communication", "formulation problem", "collection data",
           "cleaning data", "eda", "modeling")
nodes <- data.frame(
  id = 1:150,
  group = sample(steps, 150, replace = TRUE, prob = c(0.2, 0.1, 0.15, 0.15, 0.15, 0.25))
)
nodes <- nodes[order(nodes$group), ]   # the blocks of the circle follow the groups

# 100 edges: 75 join two nodes of one group, 25 join any two nodes
from <- sample(nodes$id, 100, replace = TRUE)
to <- sample(nodes$id, 100, replace = TRUE)
for (i in 1:75) {
  members <- nodes$id[nodes$group == nodes$group[nodes$id == from[i]]]
  to[i] <- members[sample.int(length(members), 1)]
}

graph <- igraph::graph_from_data_frame(data.frame(from, to), vertices = nodes, directed = FALSE)
igraph::V(graph)$degree <- igraph::degree(graph)
layout <- create_layout(graph, layout = "linear", circular = TRUE)

# a label for each group, outside the circle, at the middle of its block
angle <- tapply(seq_len(nrow(layout)), layout$group, function(i) {
  atan2(mean(layout$y[i]), mean(layout$x[i]))
})
group_labels <- data.frame(group = names(angle), x = 1.18 * cos(angle), y = 1.18 * sin(angle))
group_labels$hjust <- ifelse(group_labels$x > 0.15, 0, ifelse(group_labels$x < -0.15, 1, 0.5))

ggraph(layout) +
  geom_edge_arc(edge_colour = "sienna", edge_alpha = 0.3, edge_width = 0.4) +
  geom_node_point(aes(color = group, size = degree), alpha = 0.85) +
  geom_text(data = group_labels, aes(x, y, label = group, color = group, hjust = hjust),
            fontface = "bold", size = 3.4, show.legend = FALSE) +
  scale_color_set("npg") +
  scale_size_continuous(range = c(1, 5)) +
  coord_fixed(xlim = c(-1.75, 1.75), ylim = c(-1.35, 1.35), clip = "off") +
  theme_void() +
  theme(legend.position = "none")
```

- `layout = "linear", circular = TRUE` places the nodes on a circle in the order
  of the data, so sorting `nodes` by group is what makes the blocks.
- The labels are computed from the positions of the nodes, at the middle of each
  block, so they follow the data when the groups change.
- The colors come from `scale_color_set("npg")`. The nodes and the labels share
  the color aesthetic, so another name from `list_color_scales()` recolors both.

---

## 2. A Circos Plot

A circos plot lays tracks of genomic data around a circle. This one has the
first eight human chromosomes, whose ideogram comes with circlize, and three
tracks: bars, a line with points, and blocks that run from a loss to a gain.
Ten links cross the circle, as translocations do.

circlize draws with base graphics, so a ggplot2 theme does not apply to it.
The colors can still come from `themeset`, which keeps the figure in step with
the others of a paper.

```{r circos, eval = has_circos, fig.width = 6.5, fig.height = 6.5, out.width = "80%", fig.cap = "A circos plot of eight chromosomes with three data tracks and ten links.", fig.alt = "A circle made of the eight chromosomes with their banding, with three rings inside it: bars, a line with points, and blocks colored from blue to red. Colored links cross the middle of the circle."}
library(circlize)

set.seed(7)
chromosomes <- paste0("chr", 1:8)
nature_colors <- journal_colors("nature")

circos.clear()
circos.par(track.height = 0.1, cell.padding = c(0.02, 0, 0.02, 0))
circos.initializeWithIdeogram(species = "hg19", chromosome.index = chromosomes)

# Track 1: gene density, as bars
gene_density <- generateRandomBed(nr = 500)
circos.genomicTrack(gene_density, numeric.column = 4, panel.fun = function(region, value, ...) {
  circos.genomicRect(region, abs(value), ytop.column = 1, ybottom = 0,
                     col = nature_colors[4], border = NA, ...)
}, track.height = 0.15)

# Track 2: mutation frequency, as a line with points
circos.genomicTrack(gene_density, numeric.column = 4, panel.fun = function(region, value, ...) {
  circos.genomicLines(region, value, type = "l", col = nature_colors[3], ...)
  circos.genomicPoints(region, value, cex = 0.4, pch = 16, col = nature_colors[3], ...)
}, track.height = 0.1)

# Track 3: copy number, from a loss to a gain, through white
copy_number <- generateRandomBed(nr = 100)
gain_loss <- colorRamp2(c(-1, 0, 1), c(nature_colors[4], "white", nature_colors[1]))
circos.genomicTrack(copy_number, numeric.column = 4, panel.fun = function(region, value, ...) {
  circos.genomicRect(region, value, col = gain_loss(pmax(pmin(value[[1]], 1), -1)),
                     border = NA, ...)
}, track.height = 0.05)

# Links: ten regions joined to ten others, all on chromosomes 1 to 8
chromosome_length <- read.chromInfo(species = "hg19")$chr.len[chromosomes]
random_regions <- function(n, width = 4e6) {
  chr <- sample(chromosomes, n, replace = TRUE)
  start <- vapply(chr, function(chr) sample.int(chromosome_length[[chr]] - width, 1), numeric(1))
  data.frame(chr = chr, start = start, end = start + width)
}
link_colors <- adjustcolor(nature_colors[c(1, 2, 3, 5, 6)][sample(5, 10, replace = TRUE)], alpha.f = 0.7)
circos.genomicLink(random_regions(10), random_regions(10), col = link_colors, lwd = 2)

circos.clear()
```

- circlize keeps its settings in global state: `circos.clear()` before a plot
  and after it, so that the next one starts clean.
- `generateRandomBed()` makes the simulated regions. They are drawn at random
  over the whole genome, and only the eight chromosomes of the ideogram show.
  The links are made by hand for that reason: a link needs both of its ends on a
  chromosome that is drawn.
- `journal_colors("nature")` gives the palette as colors. The gain and loss
  track runs between two of them, through white, so its two ends are the colors
  of the other tracks.

---

## 3. A Figure of Four Panels

A figure for print: four panels with tags, assembled with patchwork. Every panel
has the theme of Nature, `journal_theme("nature")`, so the text is 7 pt, and the
figure is 183 mm wide, the double column of the journal.

The colors of the correct and the incorrect group are two Okabe-Ito colors, which
readers with color-vision deficiency can tell apart. A red and a green would not
be safe for them.

```{r panels, eval = has_patchwork, fig.width = 183 / 25.4, fig.height = 4.6, fig.cap = "A four-panel figure in the theme of Nature: overlapping histograms, a scatter plot with a fitted line, horizontal bars and a line with a marked interval.", fig.alt = "Four panels tagged a to d. Panel a has two overlapping histograms of confidence, in blue and orange. Panel b is a scatter plot with a dashed line falling from left to right. Panel c has four horizontal bars of different length. Panel d is a falling line with a shaded interval and a dashed vertical line."}
library(patchwork)

set.seed(123)
confidence <- data.frame(
  Confidence = pmin(pmax(c(rnorm(1000, 0.7, 0.15), rnorm(800, 0.5, 0.15)), 0), 1),
  Status = factor(rep(c("Correct", "Incorrect"), c(1000, 800)))
)
bins <- data.frame(
  Confidence_bin = seq(0.1, 1, by = 0.1),
  Error_rate = seq(0.9, 0.1, length.out = 10) + rnorm(10, 0, 0.05)
)
models <- data.frame(
  Feature = c("Confidence", "Difficulty", "RAG score", "Embeddings"),
  AIC = c(1327, 1347, 1356, 1358)
)

okabe_ito <- journal_colors("cell")
status <- c(Correct = okabe_ito[5], Incorrect = okabe_ito[6])
nature_theme <- journal_theme("nature")

# a: two overlapping histograms
panel_a <- ggplot(confidence, aes(Confidence, fill = Status)) +
  geom_histogram(aes(y = after_stat(density)), position = "identity", bins = 30,
                 alpha = 0.6, colour = "black", linewidth = 0.2) +
  scale_fill_manual(values = status) +
  labs(x = "Confidence", y = "Density") +
  nature_theme +
  theme(legend.position = "inside", legend.position.inside = c(0.22, 0.82),
        legend.title = element_blank())

# b: points that follow a fitted line, colored by their position
panel_b <- ggplot(bins, aes(Confidence_bin, Error_rate)) +
  geom_smooth(method = "lm", formula = y ~ x, colour = "black", linetype = "dashed",
              linewidth = 0.4, se = FALSE) +
  geom_point(aes(colour = Confidence_bin), size = 1.8) +
  scale_colour_viridis_c(guide = guide_colourbar(barheight = unit(14, "mm"),
                                                 barwidth = unit(2.5, "mm"))) +
  labs(x = "Confidence bin", y = "Error rate", colour = "Bin") +
  nature_theme

# c: horizontal bars, on an axis that starts at 1300
panel_c <- ggplot(models, aes(AIC, reorder(Feature, -AIC), fill = Feature)) +
  geom_col(orientation = "y", width = 0.6) +
  coord_cartesian(xlim = c(1300, 1400)) +
  scale_fill_set("npg") +
  labs(x = "AIC", y = NULL) +
  nature_theme +
  theme(legend.position = "none")

# d: a line, with a shaded interval and two reference lines
panel_d <- ggplot(bins, aes(Confidence_bin, Error_rate)) +
  annotate("rect", xmin = 0.7, xmax = 0.9, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = okabe_ito[6]) +
  geom_vline(xintercept = 0.8, linetype = "dashed", colour = okabe_ito[6], linewidth = 0.4) +
  geom_hline(yintercept = 0.5, linetype = "dotted", linewidth = 0.4) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.4) +
  labs(x = "Phase 1 chosen confidence", y = "Abstention probability") +
  nature_theme

(panel_a | panel_b) / (panel_c | panel_d) +
  plot_annotation(tag_levels = list(journal_tags("nature", 4)))
```

The panels are separate ggplots, so `journal_check()` can read them. It is given
the list of the four, with the double column and the height of the figure:

```{r panels_check, eval = has_patchwork}
journal_check(list(panel_a, panel_b, panel_c, panel_d), "nature", column = "double", height = 117)
```

- Each panel carries the theme of the journal, and patchwork draws the tags with
  the `plot.tag` element of it: bold, 8 pt, in lowercase for Nature.
  `journal_tags("science", 4)` would give capitals.
- The axis of panel **c** starts at 1300, so the lengths of the bars do not
  start at zero. That is a choice that a figure has to state in its legend.
- The four panels draw 16 colors between them, among them the gradient of
  panel **b**, so the check does not compare them, as its note says.
- [A Multi-Panel Figure](multi-panel-figure.html) assembles panels with cowplot
  instead, and restyles the whole figure by changing the name of a set.

---

## 4. A Grouped Dot Plot

One row for each region, with a dot for each of three age groups, on a grey
strip, and the value above each dot. The age groups have an order, so their
colors are three shades of one color, from light to dark. `journal_ramp()` makes
them from a color of a journal palette.

```{r dots, fig.width = 7, fig.height = 5.2, fig.cap = "A grouped dot plot of net approval by region and age group, in three shades of one color.", fig.alt = "Five rows, one for each region, with a grey strip behind each. On each row three dots, in light, medium and dark green, show the value for three age groups. The values are written above the dots, and the light dots, the youngest group, are furthest to the right in every row."}
regions <- c("Southeast Asia", "South Asia", "Sub-Saharan Africa", "European Union",
             "Middle East\nand North Africa")
ages <- c("15 to 29", "30 to 49", "50 and over")

approval <- data.frame(
  Region = factor(rep(regions, times = 3), levels = rev(regions)),
  Age_Group = factor(rep(ages, each = 5), levels = ages),
  Net_Approval = c(58, 49, 52, 31, -14,    # 15 to 29
                   47, 32, 38, 19, -27,    # 30 to 49
                   36, 20, 28, 8, -38)     # 50 and over
)
strips <- data.frame(Region = factor(regions, levels = rev(regions)))

ggplot(approval, aes(Net_Approval, Region)) +
  # the grey strips: one for each region, not one for each dot
  geom_linerange(data = strips, aes(y = Region, xmin = -45, xmax = 70), inherit.aes = FALSE,
                 colour = "grey95", linewidth = 9) +
  geom_point(aes(colour = Age_Group), size = 4) +
  geom_text(aes(label = Net_Approval), vjust = -1.5, size = 3) +
  scale_colour_manual(values = journal_ramp("nature", 3, color = 3, light = 0.55)) +
  scale_x_continuous(limits = c(-45, 70), breaks = seq(-40, 60, 20)) +
  labs(title = "Approval Is Higher\nAmong Younger Adults",
       subtitle = "Net approval by region and age group (simulated data)",
       x = NULL, y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.title = element_text(size = 18, face = "bold", family = "serif"),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 10, colour = "grey40")
  )
```

- `journal_ramp("nature", 3, color = 3, light = 0.55)` shades the third color of
  the Nature palette. `light` is the share of white in the lightest shade; a
  smaller number gives darker dots on the grey strips.
- The strips come from their own data frame, with `inherit.aes = FALSE`. Drawn
  from the data of the dots they would be drawn three times each.
- The title has a line break (`\n`) and a serif face. `plot.title.position =
  "plot"` aligns the title and the subtitle with the whole plot, not with the
  panel.

---

## 5. A Composite Figure of Nine Panels

The kind of figure that a paper on patient-derived models ends with: nine panels
in five rows. Two are schematics, which are drawn in a vector editor, so only a
placeholder keeps their place. The others are donut charts, two small heat maps,
an oncoprint of mutations, a UMAP, a correlation heat map, ridgelines and a
scatter plot with a fitted line.

Some panels are ggplot2 plots and some are drawn by ComplexHeatmap with grid
graphics. `patchwork::wrap_elements()` turns a grid graphic into a patch, so that
patchwork lays out all nine and tags them. The figure is 183 mm wide, the double
column of Nature, and 170 mm high, its maximum height. One palette, Okabe-Ito,
colors every category, so that the figure keeps to the colors that the guide of
Nature lists.

```{r composite, eval = has_composite, fig.width = 183 / 25.4, fig.height = 170 / 25.4, fig.cap = "A composite figure of nine panels: two schematic placeholders, donut charts, two small heat maps, an oncoprint, a UMAP, a correlation heat map, ridgelines and a scatter plot.", fig.alt = "A figure of nine panels tagged a to i, in five rows. Panels a and h are dashed boxes that stand for schematics. Panel b has three donut charts. Panel c has two small heat maps. Panel d is an oncoprint of eight genes. Panel e is a scatter plot of four clusters. Panel f is a heat map of six lineages with colored strips on its top and left. Panel g has three ridgelines. Panel i is a scatter plot with a rising fitted line."}
library(patchwork)
library(ComplexHeatmap)
library(ggridges)
library(grid)

set.seed(42)
okabe_ito <- journal_colors("cell")
nature_theme <- journal_theme("nature")
text_pt <- 5.6   # the size of axis text in the theme of Nature; the heat maps use it too
tag_style <- theme(plot.tag = element_text(size = 8, face = "bold"))

lineages <- c("Colon", "Pancreas", "Ovary", "Breast", "Stomach", "Liver")
lineage_colors <- setNames(okabe_ito[seq_along(lineages)], lineages)

# a and h: placeholders for the schematics
placeholder <- function(text) {
  ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1, fill = "grey98",
             colour = "grey70", linetype = "dashed", linewidth = 0.3) +
    annotate("text", x = 0.5, y = 0.5, label = text, size = 2.5, fontface = "italic", colour = "grey30") +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_void(base_size = 7) +
    tag_style
}
fig_a <- placeholder("Schematic of the experimental workflow\n(drawn in a vector editor)")
fig_h <- placeholder("Schematic of differentiation over time\n(drawn in a vector editor)")

# b: one donut chart for each kind of data, in one plot with three facets
donuts <- data.frame(
  data = factor(rep(c("WGS", "RNA-seq", "CRISPR"), each = 6), levels = c("WGS", "RNA-seq", "CRISPR"),
                labels = c("WGS (n = 251)", "RNA-seq (n = 306)", "CRISPR (n = 147)")),
  lineage = factor(rep(lineages, 3), levels = lineages),
  n = c(35, 25, 18, 12, 6, 4,   30, 28, 20, 12, 6, 4,   40, 22, 16, 12, 6, 4)
)
donuts$fraction <- ave(donuts$n, donuts$data, FUN = function(n) n / sum(n))
donuts$ymax <- ave(donuts$fraction, donuts$data, FUN = cumsum)
donuts$ymin <- donuts$ymax - donuts$fraction

fig_b <- ggplot(donuts) +
  geom_rect(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 2.4, fill = lineage), colour = "white", linewidth = 0.2) +
  coord_polar(theta = "y") +
  xlim(c(1, 4)) +
  facet_wrap(vars(data)) +
  scale_fill_manual(values = lineage_colors, name = NULL) +
  theme_void(base_size = 7) +
  theme(strip.text = element_text(size = text_pt, face = "bold"), legend.text = element_text(size = text_pt),
        legend.key.size = unit(2.5, "mm"), panel.spacing = unit(0, "mm")) +
  tag_style

# c: two small heat maps side by side, drawn by ComplexHeatmap into one grid graphic
diverging <- circlize::colorRamp2(c(-1, 0, 1), c(okabe_ito[5], "white", okabe_ito[6]))
small_heatmap <- function(genes, title) {
  values <- matrix(runif(16, -1, 1), nrow = 4,
                   dimnames = list(genes, c("Tumor", "Organoid", "Spheroid", "Cell line")))
  Heatmap(values, col = diverging, name = title, show_heatmap_legend = FALSE,
          cluster_rows = FALSE, cluster_columns = FALSE,
          column_title = title, column_title_gp = gpar(fontsize = 7, fontface = "bold"),
          row_names_gp = gpar(fontsize = text_pt), column_names_gp = gpar(fontsize = text_pt),
          column_names_rot = 45, rect_gp = gpar(col = "white", lwd = 0.5))
}
heatmaps <- grid.grabExpr({
  pushViewport(viewport(layout = grid.layout(1, 2)))
  pushViewport(viewport(layout.pos.col = 1))
  draw(small_heatmap(c("TP53", "APC", "KRAS", "BRAF"), "Promoter c1"), newpage = FALSE)
  popViewport()
  pushViewport(viewport(layout.pos.col = 2))
  draw(small_heatmap(c("SMAD4", "PIK3CA", "PTEN", "ERBB2"), "Promoter c2"), newpage = FALSE)
  popViewport(2)
})
fig_c <- wrap_elements(full = heatmaps) + tag_style

# d: an oncoprint of 8 genes in 30 samples, without the bar plots at its edges
genes <- c("TP53", "KRAS", "APC", "SMAD4", "BRAF", "PIK3CA", "ARID1A", "PTEN")
alterations <- matrix(
  sample(c("", "MUT", "AMP", "DEL"), length(genes) * 30, replace = TRUE, prob = c(0.72, 0.15, 0.08, 0.05)),
  nrow = length(genes), dimnames = list(genes, paste0("S", 1:30))
)
alteration_colors <- c(MUT = okabe_ito[3], AMP = okabe_ito[6], DEL = okabe_ito[5])
alter_fun <- list(
  background = function(x, y, w, h) grid.rect(x, y, w, h, gp = gpar(fill = "grey94", col = "white")),
  MUT = function(x, y, w, h) grid.rect(x, y, w, h * 0.33, gp = gpar(fill = alteration_colors[["MUT"]], col = NA)),
  AMP = function(x, y, w, h) grid.rect(x, y, w, h, gp = gpar(fill = alteration_colors[["AMP"]], col = NA)),
  DEL = function(x, y, w, h) grid.rect(x, y, w, h, gp = gpar(fill = alteration_colors[["DEL"]], col = NA))
)
oncoprint <- oncoPrint(
  alterations, alter_fun = alter_fun, col = alteration_colors, show_column_names = FALSE,
  top_annotation = NULL, right_annotation = NULL,
  row_names_gp = gpar(fontsize = text_pt), pct_gp = gpar(fontsize = text_pt),
  # the labels of this legend can only be sized if `at` and `labels` are given too
  heatmap_legend_param = list(title = "Alteration", nrow = 1,
                              at = names(alteration_colors), labels = names(alteration_colors),
                              title_gp = gpar(fontsize = text_pt, fontface = "bold"),
                              labels_gp = gpar(fontsize = text_pt),
                              grid_height = unit(2, "mm"), grid_width = unit(2, "mm"))
)
# the extra padding on the left keeps the tag of the panel off the percentages
fig_d <- wrap_elements(full = grid.grabExpr(
  draw(oncoprint, heatmap_legend_side = "bottom", padding = unit(c(1, 7, 1, 1), "mm"))
)) + tag_style

# e: a UMAP of four cell types
umap <- data.frame(
  UMAP_1 = c(rnorm(120, -2, 0.8), rnorm(120, 2, 0.7), rnorm(120, 0, 0.9), rnorm(120, 3, 0.6)),
  UMAP_2 = c(rnorm(120, 2, 0.8), rnorm(120, 1, 0.7), rnorm(120, -2, 0.8), rnorm(120, -1, 0.7)),
  Cell_type = rep(c("Epithelial", "Mesenchymal", "Endothelial", "Tumor-like"), each = 120)
)
fig_e <- ggplot(umap, aes(UMAP_1, UMAP_2, colour = Cell_type)) +
  geom_point(alpha = 0.7, size = 0.6, stroke = 0) +
  scale_colour_manual(values = okabe_ito[c(2, 3, 5, 6)]) +
  guides(colour = guide_legend(override.aes = list(size = 1.5, alpha = 1))) +
  labs(colour = NULL) +
  nature_theme

# f: a correlation heat map, with the lineage colors of panel b on its top and left
correlation <- matrix(runif(36, 0.1, 0.95), nrow = 6, dimnames = list(lineages, lineages))
correlation[lower.tri(correlation)] <- t(correlation)[lower.tri(correlation)]
diag(correlation) <- 1
lineage_strip <- function(which) {
  HeatmapAnnotation(lineage = lineages, col = list(lineage = lineage_colors), which = which,
                    show_legend = FALSE, show_annotation_name = FALSE, simple_anno_size = unit(1.6, "mm"))
}
correlation_map <- Heatmap(
  correlation, name = "Correlation",
  col = circlize::colorRamp2(c(0, 0.5, 1), journal_ramp("cell", 3, color = 5, light = 0.92)),
  cluster_rows = FALSE, cluster_columns = FALSE,
  top_annotation = lineage_strip("column"), left_annotation = lineage_strip("row"),
  row_names_side = "right", row_names_gp = gpar(fontsize = text_pt), column_names_gp = gpar(fontsize = text_pt),
  rect_gp = gpar(col = "white", lwd = 0.5),
  heatmap_legend_param = list(title_gp = gpar(fontsize = text_pt, fontface = "bold"),
                              labels_gp = gpar(fontsize = text_pt),
                              grid_width = unit(2, "mm"), legend_height = unit(14, "mm"))
)
fig_f <- wrap_elements(full = grid.grabExpr(draw(correlation_map, padding = unit(c(1, 7, 1, 1), "mm")))) + tag_style

# g: ridgelines
effects <- data.frame(
  Model = factor(rep(c("Cell lines", "Organoids", "Spheroids"), each = 400),
                 levels = c("Spheroids", "Organoids", "Cell lines")),
  Effect = c(rnorm(400, -0.6, 0.65), rnorm(400, -0.3, 0.5), rnorm(400, 0.1, 0.55))
)
fig_g <- ggplot(effects, aes(Effect, Model, fill = Model)) +
  geom_density_ridges(alpha = 0.7, scale = 1.3, rel_min_height = 0.01, colour = "white", linewidth = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  scale_fill_manual(values = c("Cell lines" = okabe_ito[5], Organoids = okabe_ito[2], Spheroids = okabe_ito[1])) +
  labs(x = "Mean gene effect", y = NULL) +
  nature_theme +
  theme(legend.position = "none")

# i: a scatter plot with a fitted line
corrected <- data.frame(Raw = runif(80, 0, 10), Noise = rnorm(80, 0, 0.8))
corrected$Corrected <- corrected$Raw * 0.82 + corrected$Noise + 0.5
fig_i <- ggplot(corrected, aes(Raw, Corrected)) +
  geom_point(colour = okabe_ito[5], alpha = 0.7, size = 0.8, stroke = 0) +
  geom_smooth(method = "lm", formula = y ~ x, colour = okabe_ito[6], fill = "grey80", linewidth = 0.4) +
  labs(x = "Raw effect", y = "Corrected effect") +
  nature_theme

fig_a / (fig_b | fig_c) / (fig_d | fig_e) / (fig_f | fig_g) / (fig_h | fig_i) +
  plot_layout(heights = c(1.2, 0.9, 1.4, 1.1, 1.1)) +
  plot_annotation(tag_levels = list(journal_tags("nature", 9)))
```

`journal_check()` reads the ggplot panels, and it is given the size of the figure.
The heat maps and the oncoprint are grid graphics that it cannot read; their text
is set to the same 5.6 pt by hand, with `text_pt`.

```{r composite_check, eval = has_composite}
journal_check(list(fig_b, fig_e, fig_g, fig_i), "nature", column = "double", height = 170)
```

- `wrap_elements(full = )` takes the grid graphic that `grid.grabExpr()` captures
  from `draw()`. Patchwork tags it like a plot. The tags of the heat maps and of
  the panels drawn with `theme_void()` come out larger and not bold, unless
  `tag_style` (bold, 8 pt) is added to each of them; the panels with
  `journal_theme()` have that style already.
- The three donut charts are one ggplot with three facets: a single plot gets a
  single tag and a shared legend. `coord_polar(theta = "y")` bends the bars of
  `geom_rect()` into rings, and `xlim()` leaves the hole in the middle.
- The colors of the lineages are one named vector, `lineage_colors`. Panel **b**
  fills the rings with it and panel **f** uses the same colors for the strips on
  its edges, so a lineage keeps its color across the figure.
- The sequential scale of panel **f** is three shades of one Okabe-Ito color, from
  `journal_ramp()`; the two ends of the scale of panel **c** are two other
  colors of the palette, through white.
- The check compares the colors of each panel with each other, not those of
  different panels, since every panel has its own legend. The panels here share
  Okabe-Ito, which keeps its colors apart for every kind of vision.

---

## 6. Packages and Credits

`themeset` calls these packages by name, and the work is theirs. They are
suggested, so `themeset` works without them.

| Package | What it draws here | Authors |
|---------|--------------------|---------|
| ggraph | the circular network | Thomas Lin Pedersen |
| igraph | the graph of the network | Gábor Csárdi, Tamás Nepusz, Vincent Traag, Szabolcs Horvát, Fabio Zanini, Daniel Noom, Kirill Müller, David Schoch and Maëlle Salmon |
| circlize | the circos plot and its ideogram | Zuguang Gu |
| patchwork | the four-panel figure and the layout of the composite figure | Thomas Lin Pedersen |
| ComplexHeatmap | the heat maps and the oncoprint of the composite figure | Zuguang Gu |
| ggridges | the ridgelines of the composite figure | Claus O. Wilke |

The palettes are those of ggsci (`npg`) and, for the Okabe-Ito colors, of base R;
see [Comparing Figures Across Journals](journal-figures.html) for their credits.

---

## Summary

| To draw | Use |
|---------|-----|
| A circular network | `ggraph(layout)` with `create_layout(graph, "linear", circular = TRUE)`, `geom_edge_arc()` and `geom_node_point()` |
| A circos plot | `circos.initializeWithIdeogram()` and `circos.genomicTrack()` of circlize, with colors from `journal_colors()` |
| A figure of several panels | patchwork, with `journal_theme()` on each panel and tags from `journal_tags()` |
| A composite of ggplot2 and ComplexHeatmap panels | `wrap_elements(full = grid.grabExpr(draw(<heat map>)))` in a patchwork layout |
| Check the panels | `journal_check(list(a, b, c, d), "nature", column = "double", height = 170)` |
| Colors for ordered groups | `journal_ramp("nature", 3)` |
| A palette by name | `scale_color_set("npg")` |
