---
title: "Global and Custom Themes"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{Global and Custom Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while a theme set is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

```{r setup}
library(ggplot2)
library(themeset)

growth <- example_growth()
od <- subset(growth, measure == "OD600")
taxa <- example_taxa()
```

## Overview

This article covers two things:

1. **Global themes.** `set_global_theme()` makes a set the default for every
   plot in the R session, so you choose the look once instead of on each plot.
2. **Custom themes.** `theme_nyt()`, `theme_midnight()`, `theme_royal()` and
   `theme_deepblue()` are themes that `themeset` defines itself.

The examples use the simulated data of `example_growth()` and
`example_taxa()`.

---

## 1. Setting a Global Theme

`set_global_theme()` takes the name of a set and does two things:

1. It calls `ggplot2::theme_set()` with the set's theme.
2. If the set has color and fill scales, it stores the first three colors of
   each palette as ggplot2's default discrete colors and fills (the options
   `ggplot2.discrete.colour` and `ggplot2.discrete.fill`).

It prints a message and returns the set invisibly. From then on plots need no
theme and no scale of their own:

```{r global_on}
set_global_theme("avatar")
```

```{r global_plots, fig.width = 8, fig.height = 3.8}
points <- ggplot(subset(taxa, taxon == "Bacteroides"),
                 aes(condition, abundance, color = condition)) +
  geom_jitter(width = 0.1, size = 2) +
  labs(x = NULL, y = "Relative abundance (%)", color = NULL)

bars <- ggplot(taxa, aes(taxon, abundance, fill = condition)) +
  stat_summary(fun = mean, geom = "col", position = position_dodge()) +
  guides(x = guide_axis(angle = 40)) +
  labs(x = NULL, y = "Mean abundance (%)", fill = NULL)

cowplot::plot_grid(points, bars, ncol = 2)
```

The colors are stored in the options, where you can look at them:

```{r global_options}
getOption("ggplot2.discrete.colour")
getOption("ggplot2.discrete.fill")
```

### The Global Palette Has Three Colors

Because only three colors are stored, a plot with up to three groups uses the
set's colors. A plot with more groups falls back to ggplot2's own default
colors for all of its groups. The three conditions below get the colors of the
set, the six strains do not:

```{r global_limit, fig.width = 8, fig.height = 3.8}
three <- ggplot(subset(taxa, taxon == "Bacteroides"),
                aes(condition, abundance, color = condition)) +
  geom_jitter(width = 0.1, size = 2, show.legend = FALSE) +
  labs(title = "3 groups", x = NULL, y = "Relative abundance (%)")

six <- ggplot(od, aes(time, value, color = strain)) +
  geom_line(show.legend = FALSE) +
  geom_point(size = 1.5, show.legend = FALSE) +
  labs(title = "6 groups", x = "Incubation time (hr)", y = "OD600")

cowplot::plot_grid(three, six, ncol = 2)
```

For plots with many groups add a scale of your own, for example
`scale_color_avatar()`, which has eight colors.

### Switching Sets

`set_global_theme()` only touches the palette options when the new set has
scales. After a switch to a theme-only set the palette of the previous set is
still in place:

```{r global_switch}
set_global_theme("minimal")
getOption("ggplot2.discrete.colour")
```

If that is not what you want, call `reset_global_theme()` between the two.

### Resetting

`reset_global_theme()` goes back to ggplot2's defaults: `theme_gray()` and no
stored palettes. It does not restore whatever theme was active before you
called `set_global_theme()`.

```{r global_reset}
reset_global_theme()
getOption("ggplot2.discrete.colour")
```

### In Scripts and Documents

Set the theme once, near the top of the script or in the setup chunk of an R
Markdown document:

```r
library(ggplot2)
library(themeset)

set_global_theme("fivethirtyeight")

# ... every plot below uses the set ...
```

---

## 2. Custom Themes

### `theme_nyt()`

A minimal theme with bold titles, a gray subtitle and dashed gridlines. The
arguments `gridline_x` and `gridline_y` switch the major gridlines of each
axis on (the default) or off:

```{r nyt, fig.width = 8, fig.height = 3.8}
substrate <- subset(growth, measure == "Substrate (mM)" &
                      strain %in% c("Strain A", "No bacteria"))

nyt_base <- ggplot(substrate, aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 2) +
  scale_color_jama() +
  labs(
    title = "Substrate use",
    subtitle = "Concentration over 67 hours",
    x = "Incubation time (hr)", y = "Substrate (mM)", color = NULL
  )

cowplot::plot_grid(
  nyt_base + theme_nyt(),
  nyt_base + labs(title = "No vertical gridlines") + theme_nyt(gridline_x = FALSE),
  ncol = 2
)
```

### Dark Themes

`theme_midnight()`, `theme_royal()` and `theme_deepblue()` color the plot
background, the panel, the text and the gridlines. They take `base_size` and
`base_family`, which are passed on to `theme_gray()`. Dark backgrounds call for
a bright palette, so these plots use `scale_color_npg()`:

```{r dark, fig.width = 8, fig.height = 7}
dark_base <- ggplot(subset(od, strain %in% c("Strain A", "Strain B", "No bacteria")),
                    aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 2) +
  scale_color_npg() +
  labs(subtitle = "Optical density", x = "Incubation time (hr)", y = "OD600", color = NULL)

cowplot::plot_grid(
  dark_base + ggtitle("theme_midnight()") + theme_midnight(),
  dark_base + ggtitle("theme_royal()") + theme_royal(),
  dark_base + ggtitle("theme_deepblue()") + theme_deepblue(),
  dark_base + ggtitle("theme_midnight(base_size = 16)") + theme_midnight(base_size = 16),
  ncol = 2
)
```

These four themes are ordinary ggplot2 themes, so everything that works with
`theme_gray()` works with them, including adding `theme()` calls afterwards to
adjust single elements.

---

## Summary

| To do | Use |
|-------|-----|
| Set the look for the whole session | `set_global_theme("<set>")` |
| Go back to ggplot2's defaults | `reset_global_theme()` |
| Have more than three global colors | Add a `scale_color_*()` or `scale_fill_*()` to the plot |
| Light theme with dashed gridlines | `theme_nyt()`, with `gridline_x` and `gridline_y` |
| Dark theme | `theme_midnight()`, `theme_royal()` or `theme_deepblue()` |
