---
title: "Comparing Figures Across Journals"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{Comparing Figures Across Journals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7.2,
  fig.height = 4.5,
  dpi        = 192,
  out.width  = "100%",
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while a theme set is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
options(knitr.kable.NA = "")
```

```{r setup}
library(ggplot2)
library(themeset)

growth <- example_growth()
taxa <- example_taxa()
has_colorspace <- requireNamespace("colorspace", quietly = TRUE)
```

## Overview

The same results often have to fit more than one journal: a first submission to
one, a resubmission to another. Journals differ in the width of a column, in the
text sizes they accept, in the case of the panel tags and, for some, in the
colors they are known for. `themeset` keeps these as *journal templates*, so
that a figure can be drawn for one journal and set next to the same figure in
another.

| Function | What it does |
|----------|--------------|
| `journal_templates()`, `journal_template()` | The numbers of each journal; a template of your own |
| `journal_theme()`, `apply_journal()` | The theme of a journal, and its palette, for one plot |
| `journal_colors()`, `journal_ramp()` | The palette of a journal, and shades of one of its colors |
| `compare_journals()` | One figure in several journals, side by side and to scale |
| `compare_palettes()` | Which palettes keep their colors apart, also for readers with color-vision deficiency |
| `journal_check()` | A table of what a figure gets right and wrong for a journal |
| `journal_fix()` | Repairs what the check finds, and checks again |
| `match_style()`, `journal_audit()` | The figures of one paper: one style, the same color for a group, and a check of them all |
| `save_journal_figure()` | Saves a figure at the width of a column |

The numbers come from the author guides of the journals, which were read in
October 2026. Journals change their guides, so read the current one before you
submit; the address of each is in the `guide` field of the template, and the
last section lists them. `themeset` is not affiliated with any journal.

The examples use the simulated data of `example_growth()` and
`example_taxa()`.

---

## 1. The Templates

`journal_templates()` lists the journals. Widths are in millimetres, text sizes
in points. `single`, `onehalf` and `double` are the columns of the journal; a
blank means the journal does not have that column.

```{r templates}
templates <- journal_templates()
columns <- c("journal", "single_mm", "onehalf_mm", "double_mm", "text_pt",
             "min_text_pt", "max_text_pt", "tag", "palette")
knitr::kable(templates[, columns])
```

Seven journals give sizes, which are the first seven rows. The last six, NEJM,
The Lancet, JAMA, BMJ, the Journal of Clinical Oncology and the Frontiers
journals, redraw the figures of authors and give no size to draw at: only their
palette is specific, and the other numbers are generic (85, 114 and 174 mm wide,
7 pt text). Science counts three columns, so for it `onehalf` is the width of
two columns and `double` of three. PLOS has no columns: a figure may be 66.8 to
190.5 mm wide, and `double` is the full width.

A template is one row, as an object that every journal function accepts in place
of the name of the journal:

```{r template}
journal_template("nature")
```

Change a field to make a template of your own, or give a new name and the
fields of your journal to make one from scratch. The fields you leave out take
the generic values:

```{r custom_template}
journal_template("nature", text_pt = 6)$text_pt

mine <- journal_template("my_journal", single_mm = 80, double_mm = 165,
                         text_pt = 6, palette = "lancet")
mine
```

---

## 2. One Plot, Several Journals

A plot of the OD600 of five strains and a no-bacteria control, without a theme
or a palette:

```{r plot}
od <- subset(growth, measure == "OD600")

p <- ggplot(od, aes(time, value, color = strain)) +
  geom_errorbar(aes(ymin = value - se, ymax = value + se), width = 2) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 60, 20)) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL)
```

`apply_journal()` adds the theme and the palette of a journal to it, and
`compare_journals()` does that for several journals and sets the copies side by
side, each at the width of its column. The page is drawn to scale: a copy that
is 57 mm wide is that much narrower than one of 89 mm. `sheet_mm` is the width
that a row may fill, so that the copies wrap to a second row.

```{r single_build}
single_page <- compare_journals(p, c("nature", "science", "cell", "pnas"),
                                sheet_mm = 200, title = "One plot at the width of a single column")
single_size <- attr(single_page, "size_mm")
```

```{r single, fig.width = single_size[["width"]] / 25.4, fig.height = single_size[["height"]] / 25.4, fig.cap = "The OD600 plot in the single column of Nature, Science, Cell Press journals and PNAS.", fig.alt = "Four copies of a line plot of OD600 against time, at the widths of four journals. The one for Science is the narrowest."}
single_page
```

The body text is 7 pt in all four copies. The text of the axes and of the
legend is a little smaller: 5.6 pt in Nature and Science, and 6 pt in Cell Press
journals and PNAS, which do not accept anything below 6 pt. What differs most is
the room that the data get. Science has the narrowest column, 57 mm, so with the
legend beside the plot the panel is short of room; putting the legend above the
plot, with `theme(legend.position = "top")`, would give some of it back.

The colors come from the journal: the palette of ggsci that is named after
Nature for the first copy, the one named after Science for the second and, for
Cell Press journals and PNAS, for which `themeset` knows no palette, the
Okabe-Ito palette.

Every figure of the page is an ordinary ggplot, so what the page shows is what
`apply_journal()` gives you for one journal:

```{r apply, eval = FALSE}
apply_journal(p, "nature")
apply_journal(p, "science", palette = "okabe_ito")   # another palette than the journal's
apply_journal(p, "cell", palette = NA)               # the theme only, for a continuous scale
```

PLOS has no single column, so its copy falls back on the full width of 190.5 mm.
To compare journals at a common width, give a number instead of a column name:
`compare_journals(p, c("nature", "plos"), column = 85)`.

---

## 3. A Figure of Several Panels

A figure of several panels is laid out for the width it is drawn at, so
`compare_journals()` also takes a function. It is called once for each journal
with the template of that journal, which carries the size of the copy in
`width_mm` and `height_mm`, and it returns the figure: a ggplot, or several
assembled with `cowplot::plot_grid()`. `journal_tags()` gives the panel tags in
the case that the journal prints.

```{r panels}
figure <- function(spec) {
  tags <- journal_tags(spec, 4)
  style <- function(plot, tag, ...) apply_journal(plot, spec, ...) + labs(tag = tag)

  # a, d: the time courses, with one legend above the first
  pa <- style(p, tags[1]) +
    theme(legend.position = "top") +
    guides(color = guide_legend(nrow = 2))
  pd <- style(
    ggplot(subset(growth, measure == "Substrate (mM)"), aes(time, value, color = strain)) +
      geom_line() +
      labs(x = "Incubation time (hr)", y = "Substrate (mM)"),
    tags[4]
  ) + theme(legend.position = "none")

  # b: the taxa in the control, colored by taxon
  pb <- style(
    ggplot(subset(taxa, condition == "Control"), aes(taxon, abundance, fill = taxon)) +
      geom_point(position = position_jitter(width = 0.15, height = 0, seed = 1),
                 shape = 21, color = "black", stroke = 0.2) +
      guides(x = guide_axis(angle = 45)) +
      labs(x = NULL, y = "Abundance (%)"),
    tags[2]
  ) + theme(legend.position = "none")

  # c: a heat map, whose continuous scale is the journal's first color
  heat <- aggregate(abundance ~ taxon + condition, taxa, mean)
  pc <- style(
    ggplot(heat, aes(condition, taxon, fill = abundance)) +
      geom_tile() +
      scale_fill_gradient(low = "white", high = journal_colors(spec, 1)) +
      labs(x = NULL, y = NULL, fill = "Mean (%)"),
    tags[3], palette = NA
  )

  # two rows, each aligned on its own: aligning the four panels as one group would
  # give every panel the margins of the widest label and the tallest legend
  top <- cowplot::plot_grid(pa, pb, nrow = 1, align = "h", axis = "tb")
  bottom <- cowplot::plot_grid(pc, pd, nrow = 1, align = "h", axis = "tb")
  cowplot::plot_grid(top, bottom, ncol = 1)
}
```

Two journals at their double column, with a height of 0.6 times the width:

```{r panels_build}
double_page <- compare_journals(figure, c("nature", "plos"), column = "double",
                                aspect = 0.6, ncol = 1,
                                title = "A four-panel figure at the width of two columns")
double_size <- attr(double_page, "size_mm")
```

```{r double, fig.width = double_size[["width"]] / 25.4, fig.height = double_size[["height"]] / 25.4, fig.cap = "The same four panels for Nature and PLOS.", fig.alt = "Two copies of a figure with four panels, one above the other. The panel tags are lowercase in the first and uppercase in the second."}
double_page
```

The function runs once for each journal, so each copy has the tags, the text
sizes and the palette of its journal. The two figures differ in their tags (a, b,
c and d for Nature, A, B, C and D for PLOS), in the text (7 pt for Nature, and
8 pt for PLOS, which asks for 8 to 12 pt) and in the palette (the one named after
Nature, and Okabe-Ito). The width is in `spec$width_mm`, so a function can also
change the layout with it, for example one column of panels below 120 mm.

Two things in `figure()` are worth copying. The panels are assembled in two rows,
each aligned on its own, and the heat map takes `palette = NA`, which adds the
theme of the journal and leaves its continuous color scale alone.

---

## 4. Colors

`journal_colors()` returns the palette of a journal, as many colors as you ask
for:

```{r colors}
journal_colors("nature", 4)
journal_colors("science", 4)
journal_colors("cell", 4)
```

The guide of Nature lists the Okabe-Ito colors as its palette for color-blind
readers. The other palettes are the ones of ggsci that take their inspiration
from a journal, which does not mean that the journal asks for them. Cell Press
journals, PNAS, PLOS and Elsevier get Okabe-Ito, since `themeset` knows no
palette for them.

```{r swatches_build, eval = has_colorspace}
journals <- c("nature", "science", "cell", "nejm", "lancet", "jama", "bmj", "jco", "frontiers")
swatches <- do.call(rbind, lapply(journals, function(journal) {
  colors <- journal_colors(journal, 6)
  data.frame(
    journal = factor(journal, levels = rev(journals)),
    position = 1:6,
    normal = colors,
    deuteranopia = colorspace::deutan(colors)
  )
}))
swatches <- rbind(
  transform(swatches, vision = "Normal vision", color = normal),
  transform(swatches, vision = "Simulated deuteranopia", color = deuteranopia)
)
```

The first six colors of each palette, as they look to a reader with normal
vision and as they look to a reader with deuteranopia, the most common kind of
color-vision deficiency, as simulated by the colorspace package:

```{r swatches, eval = has_colorspace, fig.width = 7.2, fig.height = 3.6, fig.cap = "The first six colors of the palette of each journal, with normal vision and simulated deuteranopia.", fig.alt = "Nine rows of six colored squares in two columns. In the right-hand column, which simulates deuteranopia, the squares of several palettes look alike."}
ggplot(swatches, aes(position, journal, fill = color)) +
  geom_tile(color = "white", linewidth = 0.8) +
  scale_fill_identity() +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  facet_wrap(vars(vision)) +
  labs(x = NULL, y = NULL) +
  journal_theme("nature") +
  theme(axis.line = element_blank(), axis.ticks = element_blank())
```

`compare_palettes()` puts a number on it. For each palette it gives the smallest
distance between two of its colors, in the CIEDE2000 measure of the difference
between colors, with normal vision and with each kind of deficiency. A distance
under about 10 means that two colors are easy to confuse; this is a rule of
thumb and not a standard.

```{r compare_palettes, eval = has_colorspace}
knitr::kable(compare_palettes(n = 6), digits = 1)
```

At six colors only two palettes stay above that threshold for every kind of
vision, Okabe-Ito, the one that Nature lists, and `jco`. With more colors the
distances get smaller. If a palette is flagged and the figure has to keep it,
change the figure in another way: add direct labels, or give the groups
different line types or point shapes, so that color is not the only difference.

### Groups That Have an Order

The colors of a categorical palette are for groups that differ; they give the
reader no sense of which end is high. For doses, stages or time points,
`journal_ramp()` gives shades of one color of the journal, from a light tint to
the color itself:

```{r ramp, fig.width = 5, fig.height = 3, fig.cap = "Four shades of the first color of Nature.", fig.alt = "Four bars of increasing height, in four shades of red from light to dark."}
doses <- data.frame(dose = factor(c("1", "10", "100", "1000"), levels = c("1", "10", "100", "1000")),
                    response = c(2, 3.5, 5.5, 8))

ggplot(doses, aes(dose, response, fill = dose)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = journal_ramp("nature", 4), guide = "none") +
  labs(x = "Dose (mg/kg)", y = "Response") +
  journal_theme("nature")
```

`color` picks another color of the palette (`journal_ramp("nature", 4, color = 2)`)
and `light` sets how much white is in the lightest shade.

### Colors That Stay with a Group

A group should have the same color in every figure of a paper. Name the colors
once, and use them in each figure:

```{r roles, eval = FALSE}
group_colors <- setNames(journal_colors("nature", 3), c("Control", "Diet A", "Diet B"))

ggplot(taxa, aes(taxon, abundance, fill = condition)) +
  geom_boxplot() +
  scale_fill_manual(values = group_colors) +
  journal_theme("nature")
```

---

## 5. Checking a Figure

`journal_check()` reads a plot and its theme and reports, for one journal, the
smallest and the largest text, the thinnest line, the case of the panel tags,
the height if you give it, and whether the colors can be told apart. The default
theme of ggplot2 has text that is too large for Nature:

```{r check_default}
journal_check(p, "nature")
```

With the theme of the journal, the text and the lines pass. The colors do not
quite: two of the six colors of the palette named after Nature come close enough,
for a reader with deuteranopia, to be confused.

```{r check_themed}
journal_check(apply_journal(p, "nature"), "nature", height = 60)
```

The note says what to do. With the Okabe-Ito palette everything passes:

```{r check_okabe}
journal_check(apply_journal(p, "nature", palette = "okabe_ito"), "nature", height = 60)
```

The check is a table, a data frame with one row for each check, so you can keep
it next to the figure as a record of how the figure was made:

```{r check_record, eval = FALSE}
write.csv(journal_check(fig, "nature", height = 60), "figure1-check.csv", row.names = FALSE)
```

The check reads the plot, not the picture. It does not look for text that
overlaps or panels that are out of line: look at the figure for those. For a
figure of several panels, pass a list: `journal_check(list(a, b, c, d), "nature")`
reports the worst value over the panels.

---

## 6. Repairing a Figure

`journal_fix()` changes what `journal_check()` finds wrong, checks the figure
again, and stops when the checks pass or nothing more can be changed. The
changes are small on purpose: a text that is too small is raised to the
smallest size that the journal allows and no further, a line that is too thin
gets 0.25 pt, a panel tag gets the case of the journal, and the rest of the plot
stays as it is. Take a plot that was drawn for a slide, with small text, hairline
error bars and a capital tag:

```{r fix_slide}
slide <- ggplot(od, aes(time, value, color = strain)) +
  geom_errorbar(aes(ymin = value - se, ymax = value + se), width = 2, linewidth = 0.05) +
  geom_line() +
  geom_point() +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL, tag = "A") +
  theme_classic(base_size = 5)

journal_check(slide, "nature")
```

```{r fix}
fixed <- journal_fix(slide, "nature")
attr(fixed, "fixes")
attr(fixed, "check")
```

The colors are replaced by the Okabe-Ito palette only when two of them are hard
to tell apart, and only if the new colors are easier to tell apart than the old
ones; with more than eight groups they are left alone. `fix = c("text", "lines")`
limits the repair to some of the checks. The height cannot be repaired from the
plot: give it to `save_journal_figure()`.

`journal_fix()` is for a plot whose style you want to keep. To give a plot the
whole style of the journal, `apply_journal()` is simpler.

---

## 7. The Figures of One Paper

The figures of a paper should agree with each other: a strain that is blue in
Figure 1 should be blue in Figure 2 as well, and the text of the axes should have
one size throughout. `journal_audit()` checks a list of figures. Each figure is
checked against the journal, as with `journal_check()`, and then the figures are
compared:

```{r audit}
final <- ggplot(subset(od, time == max(time)), aes(strain, value, fill = strain)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = NULL, y = "Final OD600", fill = NULL)

figures <- list(
  "Figure 1" = apply_journal(p, "nature", palette = "okabe_ito"),
  "Figure 2" = apply_journal(final, "nature", palette = "npg")
)
audit <- journal_audit(figures, "nature")
audit[, c("figure", "check", "status", "value")]
```

The two figures give the strains different colors. `match_style()` draws a plot
in the style of a reference: the theme of the reference, and its color for each
group that the two share. A group that the reference does not have gets a color
of its palette that no group of the reference uses.

```{r match}
figures[["Figure 2"]] <- match_style(final, figures[["Figure 1"]])
audit <- journal_audit(figures, "nature")
audit[, c("figure", "check", "status", "value")]
```

```{r match_plot, fig.width = 89 / 25.4, fig.height = 60 / 25.4, out.width = "50%", fig.alt = "A bar chart of the final OD600 of each strain, in the theme of Figure 1 and with the color of each strain in Figure 1."}
figures[["Figure 2"]]
```

The attribute `colors` of the audit lists the color of every group in every
figure, for a legend or a record. The article
[A Figure Set for a Reanalysis](figure-set.html) audits seven figures that share
one color contract.

---

## 8. Saving at the Size of the Journal

A text size of 7 pt is 7 pt on the page only if the figure is saved at the size
it was drawn for. `save_journal_figure()` saves at the width of a column of the
journal, in the format of the file extension. It warns if the height is more than
the journal allows. A vector format (`.pdf`) keeps the text sharp at any zoom;
use a bitmap where the journal asks for one.

```{r save, eval = FALSE}
fig <- apply_journal(p, "nature")

save_journal_figure(fig, "figure1.pdf", journal = "nature", column = "single")
save_journal_figure(fig, "figure1.png", journal = "nature", column = "double", height = 90, dpi = 600)

# the page of compare_journals() is saved at the size of the page
save_journal_figure(single_page, "comparison.png", dpi = 150)
```

Many journals ask for Arial or Helvetica. `journal_theme("nature", base_family = "Arial")`
sets the family when the font is installed on your system.

---

## 9. Sources and Credits

The numbers of the templates come from the author guides below, read in October
2026. Where a guide gives no number, the field is blank and a generic value
is used; the panel tags of every journal but Nature and Science follow
convention (the size of the text plus one point) rather than a stated size.

```{r sources}
sized <- templates[templates$sizes, ]
knitr::kable(data.frame(Journal = sized$name, Guide = sized$guide))
```

The palettes of the other journals are the palettes of the ggsci package, which
take their inspiration from them (NEJM, The Lancet, JAMA, BMJ, JCO and
Frontiers).

`themeset` calls these packages by name; the work is theirs:

| Package | What `themeset` uses it for | Authors |
|---------|-----------------------------|---------|
| ggsci | the palettes named after journals | Nan Xiao and contributors |
| grDevices (base R) | the Okabe-Ito palette | R Core Team; the palette is by Masataka Okabe and Kei Ito |
| farver | color distances in CIEDE2000 | Thomas Lin Pedersen, Berendea Nicolae and Romain Fran&ccedil;ois |
| colorspace (optional) | the simulation of color-vision deficiency | Ross Ihaka, Paul Murrell, Kurt Hornik, Jason C. Fisher, Reto Stauffer, Claus O. Wilke, Claire D. McWhite and Achim Zeileis |
| cowplot | the page of `compare_journals()` | Claus O. Wilke |
| ragg (optional) | bitmaps of `save_journal_figure()` | Thomas Lin Pedersen and others |

---

## Summary

| To | Use |
|----|-----|
| See the numbers of a journal | `journal_templates()`, `journal_template("nature")` |
| Make a template of your own | `journal_template("my_journal", single_mm = 80, double_mm = 165)` |
| Style one plot for a journal | `apply_journal(p, "nature")` or `p + journal_theme("nature")` |
| Compare a plot in several journals | `compare_journals(p, c("nature", "science", "cell"))` |
| Compare a multi-panel figure | `compare_journals(function(spec) <figure>, ...)` |
| Compare at one width | `compare_journals(p, journals, column = 85)` |
| Get the colors of a journal | `journal_colors("nature", 5)` |
| Shades for ordered groups | `journal_ramp("nature", 4)` |
| See if a palette is safe | `compare_palettes()` |
| Check a figure against a journal | `journal_check(p, "nature", height = 60)` |
| Repair what the check finds | `journal_fix(p, "nature")` |
| Give a figure the style and colors of another | `match_style(p, reference)` |
| Check all the figures of a paper | `journal_audit(list(fig1, fig2, fig3), "nature")` |
| Save at the right size | `save_journal_figure(fig, "figure1.pdf", journal = "nature")` |
