---
title: "A Figure Set for a Reporting Study"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{A Figure Set for a Reporting Study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 183 / 25.4,
  fig.height = 4.5,
  dpi        = 160,
  out.width  = "100%",
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while a theme set is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

```{r setup}
library(ggplot2)
library(themeset)

# patchwork lays out the panels of every figure; without it the code is shown but not run
has_patchwork <- requireNamespace("patchwork", quietly = TRUE)
```

## Overview

A study of spontaneous reports asks whether a drug is reported with an event more
often than the other drugs are. Its figures follow the same path every time: the
flow of the cases into the analysis, the reports of each drug, the profile of the
events of each drug, the contrast between drugs, and a few estimates that are
checked in other ways.

This article draws five such figures from simulated data. Nothing in them is a
result: the drugs are Drug 1 to Drug 11, the events fall in Category 1 to Category
8, and the counts are made up. The reporting odds ratios (ROR) are computed from
the counts, as they would be from real ones. As in
[A Figure Set for a Reanalysis](figure-set.html), the figures share a **color
contract** and the style of a journal, `journal_theme("nature")`, and
`journal_audit()` checks the set at the end. The figures here are of other kinds:
a flow chart, a bubble matrix, a heat map with a tree, and forest plots.

---

## 1. The Study and the Color Contract

For one drug and one category of events, the reports make a 2 x 2 table: *a*
reports of the drug with the event and *b* without, *c* reports of all the other
drugs with the event and *d* without. The ROR is (*a* / *b*) / (*c* / *d*), and its
95% confidence interval is exp(log ROR +/- 1.96 *s*), with *s* the square root
of 1/*a* + 1/*b* + 1/*c* + 1/*d*. A ROR above 1 means that the event is reported
more often with the drug than with the others, which is a signal, and not a risk.

Every figure takes its colors from the contract:

- **Class**: Drugs 1 to 3 are Class A, a bluish green; Drugs 4 to 7 are Class B,
  an orange; Drugs 8 to 10 are Class C, a reddish purple; Drug 11 is in no class,
  a grey. The three colors are Okabe-Ito colors that readers with color-vision
  deficiency can tell apart. Every figure uses them.
- **Direction**: a ROR below 1 is blue and a ROR above 1 is red, on a ramp that
  passes through white at 1. No class is blue or red. Figures 3 and 4 use it.
- **Evidence**: a filled circle when the 95% CI excludes 1, an open circle when it
  includes 1; in a heat map, bold numbers mean the same.
- **Categories of events** have no color. There are eight of them, more than readers
  with color-vision deficiency can keep apart (see Section 7), so they are the labels
  of the columns and are named in the text of the figures. The routes of Figure 2
  take the Okabe-Ito colors that nothing else uses, and the ordered bins of time are
  greys.

A group keeps its color across the figures, and no figure gives one color two
meanings.

```{r contract, eval = has_patchwork}
library(patchwork)

okabe_ito <- journal_colors("cell")   # orange, sky blue, bluish green, yellow, blue, vermillion, reddish purple, black

class_levels <- c("Class A", "Class B", "Class C", "No class")
class_fill <- c(`Class A` = okabe_ito[3], `Class B` = okabe_ito[1], `Class C` = okabe_ito[7], `No class` = "grey60")
direction <- c(low = okabe_ito[5], mid = "white", high = "#B2182B")       # below 1 is blue, above 1 is red

# one theme for every panel, and the sizes that the panels share
figure_theme <- journal_theme("nature")
small_text <- 5.5 / .pt                       # annotations are 5.5 pt (geom_text sizes are in mm)
strip_text <- theme(strip.background = element_blank(), strip.text = element_text(face = "bold"))
tag_theme <- theme(plot.tag = element_text(size = 8, face = "bold"))

# the panels of a figure sit on a grid of columns: each row lists its panels and their widths
design_rows <- function(rows) {
  paste(vapply(rows, function(row) {
    paste0(vapply(names(row), function(id) strrep(id, row[[id]]), ""), collapse = "")
  }, ""), collapse = "\n")
}
tag_panels <- function(figure, n) {
  figure + plot_annotation(tag_levels = list(journal_tags("nature", n)), theme = tag_theme)
}
minus <- function(x) sub("-", "\u2212", x, fixed = TRUE)      # a real minus sign
```

The study itself is made of the counts of reports. Eleven drugs have between 2,300
and 46,000 reports, and 900,000 reports of other drugs are the comparator. Each
class has a few real effects, set in `lfc`, the log of the ratio of the odds to
the comparator, and the rest is noise. Every figure reads from the table `cells`,
which has one row for each drug and category.

```{r study, eval = has_patchwork}
set.seed(101)

drugs <- sprintf("Drug %d", 1:11)
drug_group <- factor(c(rep(class_levels[1:3], times = c(3, 4, 3)), "No class"), class_levels)
names(drug_group) <- drugs
events <- sprintf("Category %d", 1:8)

n_reports <- c(46000, 21000, 8300, 31000, 14500, 5400, 2300, 18500, 6900, 3100, 9600)   # reports of each drug
n_other <- 900000                                                                       # reports of all the other drugs
p_other <- c(0.060, 0.080, 0.115, 0.060, 0.095, 0.065, 0.050, 0.045)                     # the share of those with each category

# the log of the ratio of a drug to the comparator, for each category: noise and a few effects by class
lfc <- matrix(rnorm(11 * 8, 0, 0.28), 11, 8, dimnames = list(drugs, events))
lfc[1:3, 1] <- lfc[1:3, 1] + c(1.7, 1.2, 0.8)
lfc[4:7, c(3, 6)] <- lfc[4:7, c(3, 6)] + rep(c(1.1, 0.8, 0.6, 0.4), 2)
lfc[8:10, 5] <- lfc[8:10, 5] + c(1.2, 0.9, 0.6)
lfc[8:10, 2] <- lfc[8:10, 2] - 0.6
lfc[11, c(4, 7)] <- lfc[11, c(4, 7)] + c(1.0, 0.6)

cells <- expand.grid(drug = drugs, event = events, stringsAsFactors = FALSE)
cells$n_drug <- n_reports[match(cells$drug, drugs)]
p_cell <- pmin(p_other[match(cells$event, events)] * exp(lfc[cbind(cells$drug, cells$event)]), 0.9)
cells$a <- pmax(rbinom(nrow(cells), cells$n_drug, p_cell), 1)                  # reports of the drug with the event
cells$b <- cells$n_drug - cells$a                                              # and without
cells$c <- round(n_other * p_other[match(cells$event, events)])                # reports of the others with the event
cells$d <- n_other - cells$c                                                   # and without

# the reporting odds ratio and its 95% CI, from the four counts
from_counts <- function(a, b, c, d) {
  ror <- (a / b) / (c / d)
  se <- sqrt(1 / a + 1 / b + 1 / c + 1 / d)
  data.frame(ror = ror, se = se, lo = exp(log(ror) - 1.96 * se), hi = exp(log(ror) + 1.96 * se))
}
cells <- cbind(cells, from_counts(cells$a, cells$b, cells$c, cells$d))
cells$signal <- cells$a >= 3 & cells$lo > 1 & cells$ror >= 2                   # the rule of a signal: at least 3 reports, CI above 1, ROR of 2 or more
cells$group <- drug_group[cells$drug]
cells$drug <- factor(cells$drug, drugs)
cells$event <- factor(cells$event, events)

head(cells[cells$signal, c("drug", "event", "a", "ror", "lo", "hi")], 4)
```

---

## 2. Figure 1: The Flow of the Cases

A flow chart is a plot with no axes: boxes are `geom_rect()`, their text is
`geom_text()`, and the arrows are `geom_segment()` with an `arrow()`. Everything
sits on a plane of 100 by 100, so that the position of a box is a number that
can be read and changed. The counts are computed from the study, so that the
chart agrees with the figures after it.

```{r figure_1, eval = has_patchwork, fig.width = 120 / 25.4, fig.height = 100 / 25.4, out.width = "66%", fig.cap = "Figure 1. The flow of the cases: records, unique cases, the analysis set, and its two groups.", fig.alt = "A flow chart of three boxes in dark to light grey, one below the other, joined by arrows with notes about what was removed at the right of each arrow. The last of them splits into two pale boxes, with a strip of three colors above the first and a grey strip above the second."}
n_set <- sum(n_reports) - 3400                          # a report with several of the drugs is counted once
n_union <- sum(n_reports[1:10]) - 3100                  # the cases of the three classes
n_only <- n_set - n_union                               # the cases of Drug 11 that are in no class
n_unique <- 888000; n_older <- 142000; n_records <- n_unique + n_older
n_local <- 6800; n_nodrug <- n_unique - n_set - n_local
big <- function(x) format(x, big.mark = ",", trim = TRUE)

# boxes on a plane of 100 by 100: where each sits, its three lines of text and its colors
boxes <- data.frame(
  xmin = c(20, 20, 20, 6, 60), xmax = c(70, 70, 70, 48, 86),
  ymin = c(84, 62, 40, 9, 9), ymax = c(98, 76, 54, 27, 27),
  head = c("Records", "Unique cases", "Analysis set", "Classes A to C", "Drug 11 only"),
  n = c(n_records, n_unique, n_set, n_union, n_only),
  note = c("files of the extract", "one record kept for each case", "Drugs 1 to 11", "Drugs 1 to 10", "not in a class"),
  stringsAsFactors = FALSE
)
boxes$fill <- c("grey25", "grey45", "grey82", "grey96", "grey96")
boxes$ink <- c("white", "white", "grey10", "grey10", "grey10")
boxes$x <- (boxes$xmin + boxes$xmax) / 2
boxes$y <- (boxes$ymin + boxes$ymax) / 2

# the arrows: down the stack, and from the analysis set into its two groups
trunk <- boxes$x[1]
down <- data.frame(x = c(trunk, trunk, boxes$x[4], boxes$x[5]), y = c(boxes$ymin[1], boxes$ymin[2], 33, 33),
                   yend = c(boxes$ymax[2], boxes$ymax[3], boxes$ymax[4], boxes$ymax[5]) + 0.2)
joint <- data.frame(x = c(trunk, boxes$x[4]), xend = c(trunk, boxes$x[5]), y = c(boxes$ymin[3], 33), yend = c(33, 33))
notes <- data.frame(
  x = trunk + 3, y = c(mean(c(boxes$ymin[1], boxes$ymax[2])), mean(c(boxes$ymin[2], boxes$ymax[3]))),
  label = c(sprintf("%s older versions removed", big(n_older)),
            sprintf("%s cases without these drugs\n%s with an excluded route", big(n_nodrug), big(n_local)))
)

# a strip of the class colors above the first group, and a grey one above the second
strips <- data.frame(xmin = c(6 + c(0, 14, 28), 60), xmax = c(6 + c(14, 28, 42), 86), ymin = 25.2, ymax = 27,
                     fill = unname(class_fill[c(1, 2, 3, 4)]))

arrowhead <- arrow(length = unit(1.6, "mm"), type = "closed")
flow <- ggplot() +
  geom_segment(data = joint, aes(x = x, xend = xend, y = y, yend = yend), linewidth = 0.4, colour = "grey30") +
  geom_segment(data = down, aes(x = x, xend = x, y = y, yend = yend), arrow = arrowhead, linewidth = 0.4, colour = "grey30") +
  geom_rect(data = boxes, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill), colour = "grey35", linewidth = 0.3) +
  geom_rect(data = strips, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill)) +
  geom_text(data = boxes, aes(x = x, y = ymax - 2.5 - 1.4 * (ymax < 30), label = head, colour = ink), size = 6.2 / .pt, fontface = "bold") +
  geom_text(data = boxes, aes(x = x, y = y, label = big(n), colour = ink), size = 7 / .pt, fontface = "bold") +
  geom_text(data = boxes, aes(x = x, y = ymin + 2.3, label = note, colour = ink), size = 5.2 / .pt) +
  geom_text(data = notes, aes(x = x, y = y, label = label), hjust = 0, size = small_text, colour = "grey20", lineheight = 0.95) +
  scale_fill_identity() + scale_colour_identity() +
  scale_x_continuous(limits = c(0, 100), expand = expansion(0)) + scale_y_continuous(limits = c(5, 100), expand = expansion(0)) +
  theme_void(base_size = 7) + tag_theme

tag_panels(flow, 1)
```

- The first three boxes are three steps of grey, darkest first, and the last two
  are pale. The only colors of the chart are the strips on top of the last two: the
  colors of the three classes for the box that holds them, the grey of "no class"
  for Drug 11.
- What is removed at each step is written beside its arrow, in a size that fits
  in the margin, and the number inside each box is the one that remains.
- Because the plot is made of data frames (`boxes`, `down`, `joint`), adding a
  step is adding a row, not moving shapes by hand.

---

## 3. Figure 2: Reports, Routes and Time

The reports of each drug over the years, as one line per drug, labeled at its end
and colored by class. Next to it, the routes of administration as 100% bars. The
cells with a signal are drawn as bubbles in a matrix of drugs and categories,
sized by their reports, and the time between the start of the drug and the event
follows, as bars and as a heat map of medians. The panels in the last row share
their rows, so that a drug is read across the three.

```{r figure_2, eval = has_patchwork, fig.height = 150 / 25.4, fig.cap = "Figure 2. Reports, routes and time: cumulative reports of each drug, the routes of administration, the cells with a signal, and the time to the event.", fig.alt = "A figure of five panels, a to e. Panel a has eleven rising lines on a logarithmic axis, colored by class, each labeled with its drug. Panel b has eleven horizontal bars in three colors. Panel c is a matrix of eleven rows and eight columns, with a few colored dots of different sizes and a number beside each. Panel d has eleven horizontal bars in three greys. Panel e is a grid of eleven rows and eight columns with a number in each tile, in shades of grey."}
set.seed(21)
route_colors <- c(`Route 1` = okabe_ito[2], `Route 2` = okabe_ito[4], `Route 3` = okabe_ito[8])
bins <- c("\u226430 d", "31\u2013180 d", ">180 d")
axis_drugs <- scale_y_discrete(limits = rev(drugs))

# a: the cumulative reports of each drug, from the year of its first report
years <- 2012:2025
launch <- c(2017, 2012, 2014, 2013, 2012, 2018, 2020, 2012, 2015, 2019, 2022)
cumulative <- do.call(rbind, lapply(seq_along(drugs), function(d) {
  yr <- years[years >= launch[d]]
  w <- exp(0.2 * (yr - launch[d]))
  data.frame(drug = drugs[d], year = yr, total = cumsum(as.vector(rmultinom(1, n_reports[d], w / sum(w)))))
}))
cumulative$group <- drug_group[cumulative$drug]
ends <- cumulative[cumulative$year == max(years), ]
ends <- ends[order(ends$total), ]
ends$y <- log10(ends$total)
for (i in seq_len(nrow(ends))[-1]) ends$y[i] <- max(ends$y[i], ends$y[i - 1] + 0.1)     # labels at least 0.1 apart on the log scale

# b: the routes of each drug
p1 <- runif(11, 0.2, 0.88); p3 <- runif(11, 0.01, 0.1)
route_mix <- data.frame(drug = rep(drugs, 3), route = factor(rep(names(route_colors), each = 11), names(route_colors)),
                        share = c(p1, 1 - p1 - p3, p3))
route_mix$group <- drug_group[route_mix$drug]

# c: the cells with a signal
signal <- cells[cells$signal, ]
signal$drug <- as.character(signal$drug)

# d, e: the time to the event, in three bins and as a median
onset_share <- t(sapply(1:11, function(d) { g <- rgamma(3, c(5, 3, 2) * runif(3, 0.5, 1.6)); g / sum(g) }))
onset <- data.frame(drug = rep(drugs, 3), bin = factor(rep(bins, each = 11), bins), share = as.vector(onset_share))
median_onset <- data.frame(drug = cells$drug, event = cells$event, days = round(exp(rnorm(nrow(cells), log(60), 0.7))))

panels_2 <- list(
  A = ggplot(cumulative, aes(year, total, group = drug, colour = group)) +
    geom_line(linewidth = 0.45) +
    geom_text(data = ends, aes(x = max(years) + 0.25, y = 10^y, label = drug), hjust = 0, size = small_text, show.legend = FALSE) +
    scale_colour_manual(values = class_fill, name = NULL, drop = FALSE) +
    scale_x_continuous(breaks = seq(2012, 2024, 4), expand = expansion(add = c(0.3, 2.2))) +
    scale_y_log10(breaks = c(1e2, 1e3, 1e4), labels = c("100", "1,000", "10,000")) +
    labs(x = "Year", y = "Cumulative reports", title = "Reports of each drug") +
    figure_theme,
  B = ggplot(route_mix, aes(share, drug, fill = route)) +
    geom_col(width = 0.78, colour = "white", linewidth = 0.2) +
    geom_point(data = data.frame(drug = drugs, group = drug_group), aes(x = -0.04, y = drug, colour = group),
               shape = 15, size = 1.8, inherit.aes = FALSE, show.legend = FALSE) +
    scale_fill_manual(values = route_colors, name = NULL) +
    scale_colour_manual(values = class_fill, guide = "none") +
    axis_drugs + scale_x_continuous(limits = c(-0.07, 1), breaks = c(0, 0.5, 1), labels = c("0", "50", "100"), expand = expansion(0)) +
    labs(x = "Reports (%)", y = NULL, title = "Route of administration") +
    figure_theme,
  C = ggplot(signal, aes(event, drug)) +
    geom_point(data = cells, aes(event, as.character(drug)), size = 0.3, colour = "grey80", inherit.aes = FALSE) +
    geom_point(aes(size = a, colour = group), alpha = 0.85, shape = 16, show.legend = c(size = TRUE, colour = FALSE)) +
    geom_text(aes(label = format(a, big.mark = ",", trim = TRUE)), nudge_x = 0.36, hjust = 0, size = 5 / .pt) +
    axis_drugs + scale_x_discrete(limits = events, expand = expansion(add = c(0.5, 0.9))) +
    scale_size_area(max_size = 5.2, breaks = c(100, 1000, 5000), labels = c("100", "1,000", "5,000"), name = "Reports") +
    scale_colour_manual(values = class_fill, guide = "none") +
    labs(x = NULL, y = NULL, title = "Cells with a signal") +
    figure_theme +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid.major = element_line(colour = "grey92", linewidth = 0.25)),
  D = ggplot(onset, aes(share, drug, fill = bin)) +
    geom_col(width = 0.78, colour = "white", linewidth = 0.2) +
    axis_drugs + scale_x_continuous(breaks = c(0, 0.5, 1), labels = c("0", "50", "100"), expand = expansion(0)) +
    scale_fill_manual(values = setNames(grey(c(0.85, 0.6, 0.3)), bins), name = "Time to event") +
    labs(x = "Reports (%)", y = NULL, title = "Time to event") +
    figure_theme + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()),
  E = ggplot(median_onset, aes(event, drug, fill = days)) +
    geom_tile(colour = "white", linewidth = 0.3) +
    geom_text(aes(label = days, colour = days > 200), size = 5 / .pt, show.legend = FALSE) +
    axis_drugs + scale_x_discrete(expand = expansion(0)) +
    scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
    scale_fill_gradient(low = "white", high = grey(0.3), name = "Median\ndays") +
    labs(x = NULL, y = NULL, title = "Median time to event") +
    figure_theme + theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text.y = element_blank(),
                         axis.ticks.y = element_blank(), axis.line = element_blank())
)
tag_panels(wrap_plots(panels_2, design = design_rows(list(c(A = 20, B = 16), c(C = 15, D = 8, E = 13))), heights = c(55, 80)) +
             plot_layout(guides = "collect"), 5)
```

- Panels **c**, **d** and **e** have the same rows, and only **c** names them. The
  grid of small grey dots in **c** marks the cells that are not signals, so that
  the matrix keeps the shape of the heat map beside it.
- Each legend is made once: `plot_layout(guides = "collect")` gathers them on the
  right, and a legend that two panels would repeat is switched off in one of them
  (`show.legend`), here the class colors of **c**, which **a** already shows.
- The labels at the end of the lines of **a** are spaced on the logarithmic scale
  by a loop, at least 0.1 apart, so that no ggrepel package is needed.
- The squares of the class colors in **b** are points drawn at a position just
  left of the bars, with `show.legend = FALSE` and a scale without a guide.

---

## 4. Figure 3: The Profile of the Events

Three views of the same table. How many categories of each drug carry a signal, as
bars. The ROR of every drug and category as a heat map, with the categories in the
order of a hierarchical clustering and the tree above them; the largest ROR of
each drug has a black frame. And, below, the largest ROR of each drug with its
95% CI on a logarithmic axis, with the name of its category. The drugs are in the
same order, in the same classes, throughout.

```{r figure_3, eval = has_patchwork, fig.height = 150 / 25.4, fig.cap = "Figure 3. The profile of the events: the signals of each drug, a clustered heat map of the reporting odds ratios, and the largest ratio of each drug with its confidence interval.", fig.alt = "A figure of three panels, a to c. Panel a has eleven short horizontal bars in four groups of rows, in the colors of the classes. Panel b is a heat map of the same rows and eight columns in blue and red, with a tree above the columns and a black frame around one tile in each row. Panel c has eleven dots on a logarithmic axis with horizontal lines through them, in the colors of the classes, and a category and a number at the right of each."}
cells$class_label <- factor(as.character(cells$group), class_levels, c("Class A", "Class B", "Class C", "No\nclass"))
by_class <- facet_grid(rows = vars(class_label), scales = "free_y", space = "free_y", switch = "y")
no_strip <- theme(strip.text.y = element_blank(), strip.text.y.left = element_blank())

# a: the number of categories of each drug that carry a signal
signals <- data.frame(drug = factor(drugs, drugs), n = as.vector(tapply(cells$signal, cells$drug, sum)))
signals$group <- drug_group[as.character(signals$drug)]
signals$class_label <- factor(as.character(signals$group), class_levels, c("Class A", "Class B", "Class C", "No\nclass"))

# b: the categories in the order of a hierarchical clustering of their log ratios, and the tree
log_ror <- tapply(log2(cells$ror), list(cells$drug, cells$event), identity)
hc <- hclust(dist(t(log_ror)), method = "average")
tree_segments <- function(hc) {
  n <- length(hc$order)
  x <- numeric(n); x[hc$order] <- seq_len(n)                  # where each leaf sits
  node_x <- numeric(n - 1)
  at <- function(i, what) if (i < 0) c(x[-i], 0)[what] else c(node_x[i], hc$height[i])[what]
  do.call(rbind, lapply(seq_len(n - 1), function(k) {
    a <- hc$merge[k, 1]; b <- hc$merge[k, 2]
    node_x[k] <<- (at(a, 1) + at(b, 1)) / 2
    h <- hc$height[k]
    data.frame(x = c(at(a, 1), at(a, 1), at(b, 1)), xend = c(at(a, 1), at(b, 1), at(b, 1)),
               y = c(at(a, 2), h, h), yend = c(h, h, at(b, 2)))
  }))
}
tree <- tree_segments(hc)
cells$event_ordered <- factor(cells$event, events[hc$order])
top <- cells[ave(cells$ror, cells$drug, FUN = function(x) x == max(x)) == 1, ]     # the largest ROR of each drug
top$label <- sprintf("%s: %.1f (%.1f\u2013%.1f)", top$event, top$ror, top$lo, top$hi)

panels_3 <- list(
  A = ggplot(signals, aes(n, drug, fill = group)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = n), hjust = -0.5, size = small_text) +
    by_class + scale_y_discrete(limits = rev) +
    scale_fill_manual(values = class_fill, guide = "none") +
    scale_x_continuous(limits = c(0, max(signals$n) + 1), breaks = 0:max(signals$n), expand = expansion(0)) +
    labs(x = "Categories with a signal", y = NULL) +
    figure_theme + strip_text +
    theme(strip.placement = "outside", strip.text.y.left = element_text(angle = 90)),
  B = ggplot(cells, aes(event_ordered, drug)) +
    geom_tile(aes(fill = log2(ror)), colour = "white", linewidth = 0.3) +
    geom_text(aes(label = sprintf("%.1f", ror), fontface = ifelse(signal, "bold", "plain")), size = 5 / .pt) +
    geom_tile(data = top, fill = NA, colour = "black", linewidth = 0.5) +
    by_class + scale_y_discrete(limits = rev) + scale_x_discrete(expand = expansion(0)) +
    scale_fill_gradient2(low = direction[["low"]], mid = "white", high = direction[["high"]], midpoint = 0, limits = c(-3, 3), oob = scales::squish,
                         labels = function(x) minus(format(x)), name = "log2 ROR") +
    labs(x = NULL, y = NULL) +
    figure_theme + strip_text + no_strip +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          axis.line = element_blank(), legend.position = "right", panel.spacing.y = unit(0.8, "mm")),
  C = ggplot(top, aes(ror, drug)) +
    geom_vline(xintercept = 1, linetype = "22", linewidth = 0.3, colour = "grey45") +
    geom_linerange(aes(xmin = lo, xmax = hi, colour = group), orientation = "y", linewidth = 0.5) +
    geom_point(aes(colour = group), size = 1.7) +
    geom_text(aes(x = 120, label = label), hjust = 1, size = small_text, colour = "grey20") +
    scale_colour_manual(values = class_fill, guide = "none") +
    by_class + scale_y_discrete(limits = rev) +
    scale_x_log10(limits = c(0.7, 120), breaks = c(1, 3, 10, 30), expand = expansion(0)) +
    labs(x = "Largest reporting odds ratio of each drug (95% CI, log scale)", y = NULL) +
    figure_theme + strip_text +
    theme(strip.placement = "outside", strip.text.y.left = element_text(angle = 0, hjust = 1)),
  S = plot_spacer(),
  T = ggplot(tree) +
    geom_segment(aes(x = x, xend = xend, y = y, yend = yend), linewidth = 0.3, colour = "grey30") +
    scale_x_continuous(limits = c(0.5, 8.5), expand = expansion(0)) + scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    theme_void()
)
tag_panels(wrap_plots(panels_3, design = design_rows(list(c(S = 12, T = 22), c(A = 12, B = 22), c(C = 34))), heights = c(9, 70, 55)), 3)
```

- The tree is drawn from the result of `hclust()`: each merge is a "U" of three
  segments, and the leaves sit at 1 to 8, which are the centers of the eight
  columns of **b**. With the horizontal expansion of **b** set to zero, the tree
  and the tiles span the same width, and patchwork lines them up.
- The panels that are not figures of their own, the spacer `S` and the tree `T`,
  have letters after the panels that are, so that the tags a, b and c go to the
  three panels: patchwork numbers the areas of a design in the order of their
  letters.
- The black frame is a second `geom_tile()` with `fill = NA` on the largest cell
  of each row, and bold numbers are the cells that pass the rule of a signal.
- Panels **a** and **b** have the same facets (`by_class`) so that their rows
  agree: **a** prints the names of the classes and **b** drops them with
  `strip.text.y`. Panel **c** has the same facets again, with the names turned to
  read horizontally.

---

## 5. Figure 4: A Contrast with a Reference Drug

How does each drug compare with a reference drug, Drug 1? The contrast is the ratio
of the two RORs; its standard error is the square root of the sum of the two
squared standard errors. A heat map has one row for each of the other ten drugs and
one column for each category, in a fixed order this time; a square of the class
color sits at the start of each row. Bold numbers are ratios whose 95% CI excludes
1 and that are at least 1.5-fold either way.

```{r figure_4, eval = has_patchwork, fig.height = 85 / 25.4, fig.cap = "Figure 4. The ratio of the reporting odds ratio of each drug to that of a reference drug, for each category.", fig.alt = "A heat map of ten rows in four groups and eight columns, in blue and red with a number in each tile, some in bold. A small square of a class color sits at the left of each row. A color bar is at the right and a legend of four class colors below it."}
reference <- "Drug 1"
ref <- cells[cells$drug == reference, c("event", "ror", "se")]
rel <- cells[cells$drug != reference, ]
rel$drug <- factor(as.character(rel$drug), setdiff(drugs, reference))
rel$ratio <- rel$ror / ref$ror[match(rel$event, ref$event)]
rel$se_ratio <- sqrt(rel$se^2 + ref$se[match(rel$event, ref$event)]^2)
rel$lo_ratio <- exp(log(rel$ratio) - 1.96 * rel$se_ratio)
rel$hi_ratio <- exp(log(rel$ratio) + 1.96 * rel$se_ratio)
rel$firm <- (rel$lo_ratio > 1 & rel$ratio >= 1.5) | (rel$hi_ratio < 1 & rel$ratio <= 1 / 1.5)

contrast <- ggplot(rel, aes(event, drug)) +
  geom_tile(aes(fill = log2(ratio)), colour = "white", linewidth = 0.3) +
  geom_text(aes(label = sprintf("%.1f", ratio), fontface = ifelse(firm, "bold", "plain")), size = 5 / .pt) +
  geom_point(aes(x = 0.2, colour = group), shape = 15, size = 3.6) +
  facet_grid(rows = vars(group), scales = "free_y", space = "free_y") +
  scale_y_discrete(limits = rev) + scale_x_discrete(position = "top", expand = expansion(add = c(0.6, 0))) +
  scale_fill_gradient2(low = direction[["low"]], mid = "white", high = direction[["high"]], midpoint = 0, limits = c(-3, 3), oob = scales::squish,
                       labels = function(x) minus(format(x)), name = "log2 ratio\nto Drug 1",
                       guide = guide_colourbar(order = 1, theme = theme(legend.key.height = unit(24, "mm"), legend.key.width = unit(2.4, "mm")))) +
  scale_colour_manual(values = class_fill, name = NULL, guide = guide_legend(order = 2, nrow = 1, override.aes = list(size = 2.6))) +
  labs(x = NULL, y = NULL, caption = "Bold: the 95% CI of the ratio excludes 1 and the ratio is at least 1.5-fold either way.") +
  figure_theme + strip_text +
  theme(strip.text.y = element_blank(), axis.line = element_blank(), axis.ticks = element_blank(), panel.spacing.y = unit(0.8, "mm"),
        axis.text.x.top = element_text(angle = 0), legend.position = "right", plot.caption = element_text(hjust = 0, size = 5.6))

tag_panels(contrast, 1)
```

- The class of a row is a point of `shape = 15` on the `colour` aesthetic, so that
  the `fill` of the tiles keeps its own scale and its own color bar: a plot has one
  fill scale, and the class does not need a second.
- The caption at the bottom says what bold means; it is part of the plot, so that
  it travels with the figure.
- The color bar and the legend of the classes are two guides with an `order`, which
  stacks them in the right margin; the length of the bar is set with `theme()`
  inside `guide_colourbar()`.

---

## 6. Figure 5: Selected Estimates

A few estimates are shown with the analyses that test them, as a forest plot. Each
group is a drug and a category, with an estimate from all the reports and from some
other analyses, each of which keeps a part of the reports and moves the odds of the
event; the ROR and its 95% CI are computed again from the counts that result. The
groups are the rows of a facet with its names on the left, the color is the class
of the drug, and the shape says whether the 95% CI excludes 1.

```{r figure_5, eval = has_patchwork, fig.height = 95 / 25.4, fig.cap = "Figure 5. Selected estimates: the reporting odds ratio of five drug and category pairs in the analysis of all the reports and in other analyses.", fig.alt = "A forest plot of ten rows in five groups, with the group names at the left in bold. Each row has a dot with a short horizontal line on a logarithmic axis and a number at the right. The dots are filled, except for one that is open and lies near the dashed line at 1. The dots are green, orange, purple or grey."}
set.seed(51)

# an analysis keeps a part of the reports of a drug, and moves the odds of the event to a given ratio
variant <- function(cell, keep, ratio) {
  n <- round(cell$n_drug * keep)
  odds <- ratio * cell$c / cell$d
  a <- max(rbinom(1, n, odds / (1 + odds)), 1)
  from_counts(a, n - a, cell$c, cell$d)
}
analyses <- data.frame(
  drug  = c("Drug 1", "Drug 4", "Drug 4", "Drug 4", "Drug 4", "Drug 5", "Drug 8", "Drug 11", "Drug 11", "Drug 11"),
  event = c("Category 1", rep("Category 6", 4), "Category 3", "Category 5", rep("Category 4", 3)),
  analysis = c("All reports", "All reports", "Route 1 only", "Route 2 only", "Early reports removed", "All reports", "All reports",
               "All reports", "Compared with Class A", "Compared with Class B"),
  keep  = c(NA, NA, 0.25, 0.08, 0.6, NA, NA, NA, 1, 1),
  ratio = c(NA, NA, 1.15, 1.02, 0.93, NA, NA, NA, 1.9, 2.6),
  stringsAsFactors = FALSE
)
estimates <- do.call(rbind, lapply(seq_len(nrow(analyses)), function(i) {
  cell <- cells[cells$drug == analyses$drug[i] & cells$event == analyses$event[i], ]
  est <- if (is.na(analyses$keep[i])) cell[c("ror", "se", "lo", "hi")] else variant(cell, analyses$keep[i], analyses$ratio[i])
  cbind(analyses[i, c("drug", "event", "analysis")], est)
}))
estimates$group <- drug_group[estimates$drug]
estimates$header <- factor(sprintf("%s \u2014 %s", estimates$drug, estimates$event), unique(sprintf("%s \u2014 %s", estimates$drug, estimates$event)))
estimates$excludes <- factor(estimates$lo > 1 | estimates$hi < 1, c(TRUE, FALSE))
estimates$row <- factor(seq_len(nrow(estimates)), rev(seq_len(nrow(estimates))))
estimates$label <- sprintf("%.1f (%.1f\u2013%.1f)", estimates$ror, estimates$lo, estimates$hi)

forest <- ggplot(estimates, aes(ror, row)) +
  geom_vline(xintercept = 1, linetype = "22", linewidth = 0.3, colour = "grey45") +
  geom_linerange(aes(xmin = lo, xmax = hi, colour = group), orientation = "y", linewidth = 0.5) +
  geom_point(aes(colour = group, shape = excludes), size = 2, stroke = 0.6) +
  geom_text(aes(x = 16, label = label), hjust = 1, size = small_text, colour = "grey20") +
  facet_grid(rows = vars(header), scales = "free_y", space = "free_y", switch = "y") +
  scale_y_discrete(labels = function(x) estimates$analysis[match(x, estimates$row)]) +
  scale_x_log10(limits = c(0.7, 16), breaks = c(1, 2, 5, 10), expand = expansion(0)) +
  scale_colour_manual(values = class_fill, guide = "none") +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 1), labels = c(`TRUE` = "95% CI excludes 1", `FALSE` = "95% CI includes 1"), name = NULL, drop = FALSE) +
  labs(x = "Reporting odds ratio (95% CI, log scale)", y = NULL) +
  figure_theme + strip_text +
  theme(strip.placement = "outside", strip.text.y.left = element_text(angle = 0, hjust = 1), panel.spacing.y = unit(1.5, "mm"),
        legend.position = "top", legend.justification = "left", panel.grid.major.y = element_line(colour = "grey94", linewidth = 0.25))

tag_panels(forest, 1)
```

- The names of the analyses are the labels of the rows, and the names of the groups
  are the strips of the facets, placed on the left with `switch = "y"` and turned
  to read horizontally. The height of each group follows its number of rows, with
  `space = "free_y"`.
- A filled and an open circle are the two values of the shape scale; the legend
  says what each means, and the open circle that lies near the dashed line at 1 is
  the analysis in which the signal is gone.
- The text at the right is part of the plot, so that the numbers stay aligned with
  their rows when the figure is resized. Its position is set in data units, on the
  logarithmic axis.

---

## 7. Checking the Set

`journal_audit()` takes the panels of each figure and checks every figure against
the journal and the figures against each other:

```{r audit, eval = has_patchwork}
figures <- list(
  "Figure 1" = list(A = flow),
  "Figure 2" = panels_2,
  "Figure 3" = panels_3[c("A", "B", "C")],
  "Figure 4" = list(A = contrast),
  "Figure 5" = list(A = forest)
)
journal_audit(figures, "nature", column = "double")
```

The spacer and the tree of Figure 3 are left out of the list: neither has scales or
text to check. What the audit says is the contract at work: the classes have one
color in every figure that shows them, the text of the axes has one size, and the
figures use one font. Two things in figures of this kind are easy to miss and the
audit names them: the numbers inside the tiles of a heat map, which fit at 4.5 pt
but are below the 5 pt that Nature asks for, and a flow chart in a void theme,
whose legend title is empty but still has the 11 pt of a default theme. A text
size of 5 pt and `theme_void(base_size = 7)` are the repairs.

Why the categories have no color: eight categories need eight colors, and the
distance that the audit uses (CIEDE2000, with color-vision deficiency simulated)
says that only the eight Okabe-Ito colors keep their closest pair 11 or more apart
for every kind of vision. The qualitative palettes of `hcl.colors()` and several of
`palette.colors()` have a closest pair of 4.3 or less for at least one kind. The
Okabe-Ito colors are taken by the classes and the direction, and a color that
means a class in one figure should not mean a category in the next, so the
categories are named, in the columns of the heat map and in the text of the forest
plot, and the audit has nothing to warn about.

---

## 8. Packages and Credits

The figures are drawn with ggplot2 and laid out with patchwork (Thomas Lin
Pedersen), which `themeset` suggests, not requires. The tree comes from `hclust()`
of base R. The class colors are Okabe-Ito colors, by Masataka Okabe and Kei Ito,
as they are in base R.

---

## Summary

| To | Use |
|----|-----|
| Draw a flow chart | `geom_rect()`, `geom_text()` and `geom_segment(arrow = )` on a plane of data frames |
| Show a matrix of counts | `geom_point(aes(size = ))` with `scale_size_area()`, on discrete axes |
| Put a tree above a heat map | `hclust()`, a function that turns its merges into segments, and the same horizontal expansion of zero |
| Keep a color for one meaning | name the colors once, and give a figure no color with two meanings |
| Show the evidence of an estimate | a filled and an open circle, or bold numbers, from one rule on the CI |
| Group the rows of a forest plot | `facet_grid(rows = , scales = "free_y", space = "free_y", switch = "y")` |
| Check the whole set | `journal_audit(list("Figure 1" = panels_1, ...), "nature", column = "double")` |
