---
title: "A Showcase of Themes and Colors"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{A Showcase of Themes and Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  dpi        = 144,
  out.width  = "100%",
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while a theme set is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

One plot of the simulated growth data, drawn with the theme sets, the palettes
and the journal templates of `themeset`. Each section shows the call that makes
it.

```{r setup}
library(ggplot2)
library(themeset)

has_patchwork <- requireNamespace("patchwork", quietly = TRUE)

od <- subset(example_growth(), measure == "OD600")
p <- ggplot(od, aes(time, value, color = strain)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.4) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL)
```

## Twelve Theme Sets

Journal styles on white, news and magazine styles with their own background, and
the three dark themes of `themeset`, which suit slides. The journal sets bring
their own palette. Most of the other sets change only the theme, and
`fivethirtyeight` has a palette of three colors, too few for six groups, so these
copies use the Okabe-Ito palette, which has a color for each group.

```{r themes, eval = has_patchwork, fig.width = 16, fig.height = 11, fig.alt = "Twelve copies of a line plot of the OD600 of six strains, each in a theme set: npg, lancet, jco, nejm, fivethirtyeight, economist, wsj, solarized, nyt, midnight, royal and deepblue."}
with_palette <- c("npg", "lancet", "jco", "nejm")
sets <- c(with_palette, "fivethirtyeight", "economist", "wsj", "solarized",
          "nyt", "midnight", "royal", "deepblue")

panels <- lapply(sets, function(set) {
  styled <- apply_theme_set(p, set)
  if (!set %in% with_palette) styled <- styled + scale_color_set("okabe_ito")
  styled +
    labs(title = set) +
    theme(legend.position = "none", plot.title = ggtext::element_markdown(size = 13, face = "bold"))
})
patchwork::wrap_plots(panels, ncol = 4)
```

One set for one plot, or for the whole session:

```{r themes_use, eval = FALSE}
apply_theme_set(p, "midnight")
set_global_theme("npg")
```

## Palettes for Readers with Color-Vision Deficiency

`compare_palettes()` gives, for each palette, the distance between its two
closest colors in CIEDE2000, with normal vision and with three simulated
deficiencies. The bars below show the worst of the four; at 10 or more the
colors are easy to tell apart.

```{r palettes, fig.width = 12, fig.height = 6.5, fig.alt = "Swatches of twelve palettes with their hex codes, sorted by the distance between their two closest colors, from okabe_ito at the top to d3 at the bottom."}
palettes <- c("okabe_ito", "npg", "aaas", "nejm", "lancet", "jama", "jco",
              "d3", "futurama", "startrek", "tron", "uchicago")
scores <- compare_palettes(palettes, n = 8)
scores <- scores[order(scores$worst), ]

swatches <- do.call(rbind, lapply(scores$set, function(set) {
  colors <- unname(scale_color_set(set)$palette(8))
  colors <- colors[!is.na(colors)][seq_len(min(8, sum(!is.na(colors))))]
  data.frame(set = set, position = seq_along(colors), color = substr(colors, 1, 7))
}))
labels <- sprintf("%s  (%.0f)", scores$set, scores$worst)
swatches$set <- factor(swatches$set, levels = scores$set, labels = labels)
light <- farver::convert_colour(farver::decode_colour(swatches$color), "rgb", "lab")[, 1] > 60

ggplot(swatches, aes(position, set, fill = color)) +
  geom_tile(width = 0.92, height = 0.8) +
  geom_text(aes(label = color), size = 2.6, color = ifelse(light, "black", "white")) +
  scale_fill_identity() +
  scale_x_continuous(breaks = NULL) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 13) +
  theme(panel.grid = element_blank())
```

```{r palettes_table}
knitr::kable(scores[order(-scores$worst), c("set", "n", "normal", "deutan", "protan", "tritan", "worst")],
             digits = 1, row.names = FALSE)
```

Any of these palettes, by name:

```{r palettes_use, eval = FALSE}
p + scale_color_set("okabe_ito")
```

## One Plot in Four Journals

`compare_journals()` draws the plot at the width of a single column of each
journal, with 7 pt text and the palette of the journal, side by side and to
scale. The narrow column of Science leaves the least room for the data.

```{r journals_build}
page <- compare_journals(p, c("nature", "science", "cell", "pnas"), sheet_mm = 400)
size <- attr(page, "size_mm")
```

```{r journals, fig.width = size[["width"]] / 25.4, fig.height = size[["height"]] / 25.4, fig.alt = "The line plot at the single-column widths of Nature, Science, Cell Press and PNAS, side by side and to scale, each in the theme and palette of its journal."}
page
```

The article [Comparing Figures Across Journals](journal-figures.html) explains
the templates, and how to check and repair a figure for a journal.
