---
title: "A Showcase of Themes and Colors"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{A Showcase of Themes and Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  dpi        = 144,
  out.width  = "100%",
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while a theme set is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

One plot of the simulated growth data, drawn with the theme sets, the palettes
and the journal templates of `themeset`. Each section shows the call that makes
it.

```{r setup}
library(ggplot2)
library(themeset)

has_patchwork <- requireNamespace("patchwork", quietly = TRUE)
has_colorspace <- requireNamespace("colorspace", quietly = TRUE)

od <- subset(example_growth(), measure == "OD600")
p <- ggplot(od, aes(time, value, color = strain)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.4) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL)
```

## Twelve Theme Sets

Journal styles on white, news and magazine styles with their own background, and
the three dark themes of `themeset`, which suit slides. The journal sets bring
their own palette. Most of the other sets change only the theme, and
`fivethirtyeight` has a palette of three colors, too few for six groups, so these
copies use the Okabe-Ito palette, which has a color for each group.

```{r themes, eval = has_patchwork, fig.width = 16, fig.height = 11, fig.alt = "Twelve copies of a line plot of the OD600 of six strains, each in a theme set: npg, lancet, jco, nejm, fivethirtyeight, economist, wsj, solarized, nyt, midnight, royal and deepblue."}
with_palette <- c("npg", "lancet", "jco", "nejm")
sets <- c(with_palette, "fivethirtyeight", "economist", "wsj", "solarized",
          "nyt", "midnight", "royal", "deepblue")

panels <- lapply(sets, function(set) {
  styled <- apply_theme_set(p, set)
  if (!set %in% with_palette) styled <- styled + scale_color_set("okabe_ito")
  styled +
    labs(title = set) +
    theme(legend.position = "none", plot.title = ggtext::element_markdown(size = 13, face = "bold"))
})
patchwork::wrap_plots(panels, ncol = 4)
```

One set for one plot, or for the whole session:

```{r themes_use, eval = FALSE}
apply_theme_set(p, "midnight")
set_global_theme("npg")
```

## Palettes for Readers with Color-Vision Deficiency

`compare_palettes()` gives, for each palette, the distance between its two
closest colors in CIEDE2000, with normal vision and with three simulated
deficiencies. The bars below show the worst of the four; at 10 or more the
colors are easy to tell apart.

```{r palettes, fig.width = 12, fig.height = 6.5, fig.alt = "Swatches of twelve palettes with their hex codes, sorted by the distance between their two closest colors, from okabe_ito at the top to d3 at the bottom."}
palettes <- c("okabe_ito", "npg", "aaas", "nejm", "lancet", "jama", "jco",
              "d3", "futurama", "startrek", "tron", "uchicago")
scores <- compare_palettes(palettes, n = 8)
scores <- scores[order(scores$worst), ]

swatches <- do.call(rbind, lapply(scores$set, function(set) {
  colors <- unname(scale_color_set(set)$palette(8))
  colors <- colors[!is.na(colors)][seq_len(min(8, sum(!is.na(colors))))]
  data.frame(set = set, position = seq_along(colors), color = substr(colors, 1, 7))
}))
labels <- sprintf("%s  (%.0f)", scores$set, scores$worst)
swatches$set <- factor(swatches$set, levels = scores$set, labels = labels)
light <- farver::convert_colour(farver::decode_colour(swatches$color), "rgb", "lab")[, 1] > 60

ggplot(swatches, aes(position, set, fill = color)) +
  geom_tile(width = 0.92, height = 0.8) +
  geom_text(aes(label = color), size = 2.6, color = ifelse(light, "black", "white")) +
  scale_fill_identity() +
  scale_x_continuous(breaks = NULL) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 13) +
  theme(panel.grid = element_blank())
```

```{r palettes_table}
knitr::kable(scores[order(-scores$worst), c("set", "n", "normal", "deutan", "protan", "tritan", "worst")],
             digits = 1, row.names = FALSE)
```

Any of these palettes, by name:

```{r palettes_use, eval = FALSE}
p + scale_color_set("okabe_ito")
```

## One Figure, Twelve Palettes

The same figure, the growth curves with their error bars in the theme of Nature,
with only the palette changed. The panels are sorted by the distance between the
two closest of the six colors that the figure uses: the score depends on the
number of groups, so a palette that ranks high for eight colors can rank lower
for six, and the other way around.

```{r swap, eval = has_patchwork, fig.width = 15, fig.height = 10.5, fig.alt = "Twelve copies of the growth figure, each with a different palette, sorted from the easiest colors to tell apart to the hardest."}
fig <- ggplot(od, aes(time, value, color = strain)) +
  geom_errorbar(aes(ymin = value - se, ymax = value + se), width = 1.5, linewidth = 0.4) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.6) +
  scale_x_continuous(breaks = seq(0, 60, 20)) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL) +
  journal_theme("nature") +
  theme(text = element_text(size = 10), axis.text = element_text(size = 8),
        legend.text = element_text(size = 8), legend.key.size = unit(3, "mm"))

six <- compare_palettes(palettes, n = 6)
six <- six[order(-six$worst), ]

swapped <- Map(function(set, score) {
  fig + scale_color_set(set) +
    labs(title = sprintf("%s   %.0f%s", set, score, if (score >= 10) "  safe" else "")) +
    theme(plot.title = element_text(size = 11, face = "bold"))
}, six$set, six$worst)
patchwork::wrap_plots(swapped, ncol = 4)
```

### As Readers with Color-Vision Deficiency See It

Three of the palettes, with normal vision and as simulated for deuteranopia,
protanopia and tritanopia with the colorspace package. With d3 and npg some
strains turn into the same olive or blue, and the reader can no longer match a
line to the legend; Okabe-Ito keeps the six lines apart.

```{r cvd, eval = has_patchwork && has_colorspace, fig.width = 15, fig.height = 9, fig.alt = "The growth figure in the okabe_ito, npg and d3 palettes, each with normal vision and with simulated deuteranopia, protanopia and tritanopia."}
simulate <- list(
  "normal vision" = identity,
  deuteranopia = colorspace::deutan,
  protanopia = colorspace::protan,
  tritanopia = colorspace::tritan
)
seen <- list()
for (set in c("okabe_ito", "npg", "d3")) {
  colors <- unname(scale_color_set(set)$palette(6))[1:6]
  for (vision in names(simulate)) {
    seen[[length(seen) + 1]] <- fig +
      scale_color_manual(values = simulate[[vision]](colors)) +
      labs(title = paste0(set, ", ", vision)) +
      theme(plot.title = element_text(size = 11, face = "bold"), legend.position = "none")
  }
}
patchwork::wrap_plots(seen, ncol = 4)
```

To score the palettes for the number of groups of your own figure, and to switch
a figure to Okabe-Ito only when its colors are too close:

```{r swap_use, eval = FALSE}
compare_palettes(c("jco", "okabe_ito", "npg", "d3"), n = 6)
journal_fix(fig + scale_color_set("d3"), "nature", fix = "colors")
```

## One Plot in Four Journals

`compare_journals()` draws the plot at the width of a single column of each
journal, with 7 pt text and the palette of the journal, side by side and to
scale. The narrow column of Science leaves the least room for the data.

```{r journals_build}
page <- compare_journals(p, c("nature", "science", "cell", "pnas"), sheet_mm = 400)
size <- attr(page, "size_mm")
```

```{r journals, fig.width = size[["width"]] / 25.4, fig.height = size[["height"]] / 25.4, fig.alt = "The line plot at the single-column widths of Nature, Science, Cell Press and PNAS, side by side and to scale, each in the theme and palette of its journal."}
page
```

The article [Comparing Figures Across Journals](journal-figures.html) explains
the templates, and how to check and repair a figure for a journal.
