---
title: "Introduction to themeset"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{Introduction to themeset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while the "olink" theme is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

```{r setup}
library(ggplot2)
library(themeset)
```

## Overview

A **theme set** is a ggplot2 theme bundled with the discrete color and fill
scales that go with it. `themeset` ships 49 of them under short names such as
`"simpsons"`, `"nejm"` or `"minimal"`, and a few functions to apply a set to one
plot or to every plot in your session.

Fourteen sets carry their own scales:

- `simpsons`, `avatar`, `olink` and `fivethirtyeight`
- eight journal sets, `npg`, `aaas`, `nejm`, `lancet`, `jama`, `jco`, `bmj` and
  `frontiers`, which pair the classic theme with a journal palette
- `flexoki_light` and `flexoki_dark`

The other 35 change only the theme.

The pieces are also exported on their own, so you can mix and match:

- discrete color and fill scales (`scale_color_npg()`, `scale_fill_jama()`, ...),
  and any palette by name with `scale_color_set("nejm")`
- markdown-enabled versions of the base ggplot2 themes (`md_theme_minimal()`, ...)
- a few custom themes (`theme_nyt()`, `theme_midnight()`, ...)
- templates of journals, with their figure widths, text sizes, panel tags and
  palettes, to compare one figure in several journals (`compare_journals()`)

This vignette is a short tour. The other articles go deeper:
[Theme Sets and Color Scales](theme-sets-and-scales.html),
[Markdown-Enabled Themes](markdown-themes.html),
[Global and Custom Themes](global-and-custom-themes.html),
[A Multi-Panel Figure](multi-panel-figure.html),
[Comparing Figures Across Journals](journal-figures.html),
[A Gallery of Figure Types](figure-gallery.html),
[A Showcase of Themes and Colors](showcase.html) and
[A Figure Set for a Reanalysis](figure-set.html).

### Example Data

The examples use two small simulated data sets that come with the package as
code, so there are no data files. `example_growth()` is a growth experiment:
five bacterial strains and an uninoculated control, followed over 67 hours and
measured four ways. `example_taxa()` compares the relative abundance of five
taxa in three conditions.

```{r data_peek}
head(example_growth(), 3)
head(example_taxa(), 3)
```

### Function Reference

| Category | Function | Description |
|----------|----------|-------------|
| **Theme sets** | `apply_theme_set()` | Add a set's theme, color scale and fill scale to one plot |
| | `set_global_theme()` | Make a set the default for every plot in the session |
| | `reset_global_theme()` | Go back to ggplot2's default theme and palettes |
| | `list_theme_sets()` | Print the names of the 49 available sets |
| **Color scales** | `scale_color_npg()`, `scale_fill_npg()` | Discrete scales; also `jama`, `avatar`, `simpsons`, `olink` and `fivethirtyeight` |
| | `scale_color_set()`, `scale_fill_set()` | Discrete scales picked by name, for example `"nejm"` or `"wesanderson::Darjeeling1"` |
| | `list_color_scales()` | Print the names of the color scale sets |
| **Themes** | `md_theme_minimal()` and seven more | Markdown-enabled versions of the base ggplot2 themes |
| | `theme_nyt()` | Minimal theme with dashed gridlines and bold titles |
| | `theme_midnight()`, `theme_royal()`, `theme_deepblue()` | Dark themes |
| **Journal figures** | `journal_templates()` | The column widths, text sizes, panel tags and palettes of 13 journals |
| | `apply_journal()`, `journal_theme()` | Add the theme and the palette of a journal to a plot |
| | `compare_journals()` | Draw one figure in several journals, side by side and to scale |
| | `journal_check()` | Check a figure against the text and line limits of a journal |
| | `save_journal_figure()` | Save a figure at the width of a column of a journal |
| | `journal_colors()`, `journal_ramp()`, `compare_palettes()` | The colors of a journal, shades for ordered groups, and how well palettes keep colors apart |
| **Example data** | `example_growth()`, `example_taxa()` | Simulated growth curves and taxon abundances |

---

## 1. Quick Start

`apply_theme_set()` takes a plot and the name of a set. The plot shows the
optical density of five strains and an uninoculated control over time:

```{r quick}
od <- subset(example_growth(), measure == "OD600")

p <- ggplot(od, aes(time, value, color = strain)) +
  geom_errorbar(aes(ymin = value - se, ymax = value + se), width = 1.5) +
  geom_line() +
  geom_point(size = 2) +
  labs(title = "Growth of five strains", x = "Incubation time (hr)", y = "OD600", color = NULL)

apply_theme_set(p, "simpsons")
```

The theme, the color scale and the fill scale of the `simpsons` set were all
added to `p`. The result is an ordinary ggplot object, so you can keep adding
layers to it or save it with `ggsave()`.

Change the name to change the whole look:

```{r quick_other}
apply_theme_set(p, "avatar")
```

A journal set pairs the classic theme, which suits a paper, with the palette of
a journal:

```{r quick_journal}
apply_theme_set(p, "nejm")
```

An unknown name is an error that points back to the list of sets:

```r
apply_theme_set(p, "nope")
#> Error: Set 'nope' not found. Run list_theme_sets() for available options.
```

---

## 2. Browsing the Available Sets

```{r list_sets}
list_theme_sets()
```

The first fourteen are the complete sets. Every other name selects a theme only.
The [Theme Sets and Color Scales](theme-sets-and-scales.html) article shows
what they look like and which package each one comes from.

The color scales can be listed too:

```{r list_scales}
list_color_scales()
```

---

## 3. Using Themes and Scales Separately

Nothing forces you to go through a set. Themes and scales are ordinary ggplot2
components and can be combined freely:

```{r separate}
taxa <- example_taxa()

ggplot(taxa, aes(taxon, abundance, fill = condition)) +
  geom_boxplot() +
  labs(x = NULL, y = "Relative abundance (%)", fill = NULL) +
  theme_nyt() +
  scale_fill_jama()
```

---

## 4. Setting a Theme for the Whole Session

`set_global_theme()` makes a set the default. Plots created afterwards use its
theme and, for sets that have scales, its colors, without any extra code:

```{r global}
set_global_theme("fivethirtyeight")

ggplot(subset(taxa, taxon == "Proteobacteria"), aes(condition, abundance, fill = condition)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title = "This plot has no theme or scale of its own", x = NULL, y = "Relative abundance (%)")
```

`reset_global_theme()` returns to ggplot2's defaults:

```{r global_reset}
reset_global_theme()
```

The [Global and Custom Themes](global-and-custom-themes.html) article covers
what exactly is changed, and the limits of the global palettes.

---

## Getting Help

```r
?apply_theme_set
?set_global_theme
?theme_color_scales
?ggplot2_md_themes
?custom_themes
?example_data
```
