---
title: "Theme Sets and Color Scales"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    includes:
      before_body: lightbox.html
vignette: >
  %\VignetteIndexEntry{Theme Sets and Color Scales}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4.5,
  message    = FALSE,
  warning    = FALSE
)
# Keep vignette builds free of font downloads: OlinkAnalyze can otherwise try to
# register a web font while the "olink" theme is being built.
options(OlinkAnalyze.allow.font.load = FALSE)
```

```{r setup}
library(ggplot2)
library(themeset)

growth <- subset(example_growth(), measure == "OD600")
taxa <- example_taxa()
```

## Overview

`apply_theme_set()` adds up to three things to a plot: a theme, a discrete
color scale and a discrete fill scale. This article looks at each of them,
using the simulated data of `example_growth()` and `example_taxa()`:

1. the 14 **complete sets**, which bring their own scales,
2. the **color and fill scales**, which you can also use on their own and pick
   by name,
3. the 35 **theme-only sets**, and where each one comes from,
4. how to **adjust** a set after applying it,
5. the **sources and credits** of everything that `themeset` reaches.

---

## 1. The Complete Sets

### TV and House Sets

`simpsons`, `avatar`, `olink` and `fivethirtyeight` are complete: applying one
changes the theme and also sets the colors of the `color` and `fill`
aesthetics. Boxplots use both, the outline for `color` and the box for `fill`,
so a single plot shows what a set does. The plot compares the three conditions
for one taxon:

```{r complete_sets, fig.width = 8, fig.height = 7}
proteo <- subset(taxa, taxon == "Proteobacteria")

base <- ggplot(proteo, aes(condition, abundance, color = condition, fill = condition)) +
  geom_boxplot(alpha = 0.5, show.legend = FALSE) +
  labs(x = NULL, y = "Relative abundance (%)")

complete <- c("simpsons", "avatar", "olink", "fivethirtyeight")
panels <- lapply(complete, function(s) apply_theme_set(base, s) + ggtitle(s))
cowplot::plot_grid(plotlist = panels, ncol = 2)
```

The scales behind these sets are exported as `scale_color_simpsons()`,
`scale_color_avatar()`, `scale_color_olink()` and
`scale_color_fivethirtyeight()`, with matching `scale_fill_*()` versions.

### Journal Sets

The journal sets pair the `classic` theme, with its white background, L-shaped
axes and no grid, with a palette that is named after a journal: `npg`, `aaas`,
`nejm`, `lancet`, `jama`, `jco`, `bmj` and `frontiers`. The palettes come from
the ggsci package. They take their inspiration from the journals, and
`themeset` is not affiliated with any of them. Each panel draws the six strains
of the growth experiment:

```{r journal_sets, fig.width = 8, fig.height = 7}
strain_lines <- ggplot(growth, aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 1.2) +
  labs(x = NULL, y = NULL, color = NULL)

journal <- c("npg", "aaas", "nejm", "lancet", "jama", "jco", "bmj", "frontiers")
panels <- lapply(journal, function(s) {
  apply_theme_set(strain_lines, s) + ggtitle(s) + theme(legend.position = "none")
})
cowplot::plot_grid(plotlist = panels, ncol = 2)
```

### Flexoki

`flexoki_light` and `flexoki_dark` use the
[Flexoki](https://stephango.com/flexoki) color scheme by Steph Ango, through
the flexoki package: warm paper and ink colors that are meant for reading. The
light set takes the deep accent tones, which read well on paper, and the dark
set the paler ones, which read well on a dark screen. The dark set also draws
the default geoms in a light color (see the dark sets below).

```{r flexoki, fig.width = 8, fig.height = 3.4}
flexoki <- lapply(c("flexoki_light", "flexoki_dark"), function(s) {
  apply_theme_set(strain_lines, s) + ggtitle(s) + theme(legend.position = "none")
})
cowplot::plot_grid(plotlist = flexoki, ncol = 2)
```

---

## 2. Color and Fill Scales

The scales of the sets are available on their own. `scale_color_set()` and
`scale_fill_set()` pick a palette by name, and `list_color_scales()` prints the
names, grouped by the package that the palettes come from:

```{r list_scales}
list_color_scales()
```

| Package | Names | Colors |
|---------|-------|--------|
| ggsci | 24 palettes, among them `npg`, `aaas`, `nejm`, `lancet`, `jama`, `jco`, `bmj`, `d3`, `igv`, `futurama` and `startrek` | 6 to 79, most of them 7 to 12 |
| tvthemes | `avatar`, `simpsons` | 8 and 10 |
| OlinkAnalyze | `olink` | no fixed limit |
| ggthemes | `fivethirtyeight` | 3 |
| flexoki | `flexoki_light`, `flexoki_dark` | 8, then they repeat |
| wesanderson, biopalette, ggpalettes | `package::palette` | see below |

The oldest scales also have a function of their own: `scale_color_npg()` and
`scale_color_set("npg")` are the same scale. `olink` interpolates between its
base colors, so it can serve any number of levels, while the two flexoki sets
repeat their eight colors. The figure shows the first 12 colors of fourteen of
the sets, in the order ggplot2 hands them to the levels of a factor:

```{r palettes, fig.width = 7.5, fig.height = 5}
# The largest number of colors, up to n, that a set can give
palette_colors <- function(set, n = 12) {
  scale <- scale_color_set(set)
  for (k in n:1) {
    colors <- tryCatch(suppressWarnings(scale$palette(k)), error = function(e) NULL)
    if (!is.null(colors) && !anyNA(colors)) return(colors)
  }
}

swatch_plot <- function(sets) {
  swatches <- do.call(rbind, lapply(sets, function(set) {
    colors <- palette_colors(set)
    data.frame(set = set, position = seq_along(colors), color = colors)
  }))
  swatches$set <- factor(swatches$set, levels = rev(sets))

  ggplot(swatches, aes(position, set, fill = color)) +
    geom_tile(color = "white", linewidth = 0.8) +
    scale_fill_identity() +
    scale_x_continuous(breaks = 1:12, expand = c(0, 0)) +
    labs(x = "Position in the palette", y = NULL) +
    theme_minimal() +
    theme(panel.grid = element_blank())
}

swatch_plot(c("npg", "aaas", "nejm", "lancet", "jama", "jco", "bmj", "d3", "avatar",
              "simpsons", "olink", "fivethirtyeight", "flexoki_light", "flexoki_dark"))
```

### Optional Palettes

Three more packages are suggested, not required: wesanderson, biopalette and
ggpalettes. When one of them is installed, its palettes appear in
`list_color_scales()` as `package::palette` and work like any other name. The
qualitative palettes of biopalette are meant for biomedical figures, and the
categorical palettes of ggpalettes interpolate, so they serve many levels:

```{r optional_palettes, eval = all(vapply(c("wesanderson", "biopalette", "ggpalettes"), requireNamespace, logical(1), quietly = TRUE)), fig.width = 7.5, fig.height = 3.2}
swatch_plot(c("wesanderson::Darjeeling1", "wesanderson::Zissou1",
              "biopalette::cancer_mosaic", "biopalette::tam_pastel",
              "ggpalettes::clinical", "ggpalettes::meadow"))
```

```{r optional_use, eval = requireNamespace("biopalette", quietly = TRUE), fig.width = 6.5, fig.height = 3.6}
ggplot(taxa, aes(taxon, abundance, fill = condition)) +
  stat_summary(fun = mean, geom = "col", position = position_dodge()) +
  labs(x = NULL, y = "Mean relative abundance (%)", fill = NULL) +
  guides(x = guide_axis(angle = 40)) +
  theme_minimal() +
  scale_fill_set("biopalette::tam_pastel")
```

A palette of a package that is not installed is an error that says which
package to install.

### Using a Scale With Any Theme

The scales are plain ggplot2 scales, so they work with any theme:

```{r scales_usage, fig.width = 8, fig.height = 4.4}
lines <- ggplot(growth, aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL) +
  theme_minimal() +
  scale_color_npg()

bars <- ggplot(taxa, aes(taxon, abundance, fill = condition)) +
  stat_summary(fun = mean, geom = "col", position = position_dodge()) +
  labs(x = NULL, y = "Mean relative abundance (%)", fill = NULL) +
  guides(x = guide_axis(angle = 40)) +
  theme_minimal() +
  scale_fill_jama()

cowplot::plot_grid(lines, bars, ncol = 2)
```

### Palette Size

A palette with a fixed number of colors cannot color more levels than it has.
`fivethirtyeight` has three, so the six strains of the growth experiment leave
three of them without a color. ggplot2 warns, and the lines and points of those
strains are dropped:

```{r palette_limit, warning = TRUE, fig.width = 6.5, fig.height = 3.8}
ggplot(growth, aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL) +
  scale_color_fivethirtyeight()
```

For data with many groups choose a larger palette such as `npg`, `simpsons` or
`olink`. The palettes of wesanderson are fixed vectors of colors, so with too
many levels they stop with an error instead.

---

## 3. Theme-Only Sets

The other 35 sets change the theme and leave the scales alone. Most of them
come from other packages; `themeset` wraps each one so that
`apply_theme_set()` can reach it by name.

| Source | Sets |
|--------|------|
| ggplot2 | `bw`, `classic`, `dark`, `light`, `linedraw`, `minimal` |
| ggthemes | `base`, `calc`, `clean`, `economist`, `economist_white`, `excel`, `excel_new`, `few`, `foundation`, `gdocs`, `hc`, `igray`, `map_gg`, `pander`, `par`, `solarized`, `solarized_2`, `solid`, `stata`, `tufte`, `wsj` |
| hrbrthemes | `ipsum`, `ipsum_rc` |
| cowplot | `cowplot`, `minimal_grid` |
| themeset | `nyt`, `midnight`, `royal`, `deepblue` |

(`fivethirtyeight` is also a ggthemes theme; it is one of the complete sets
above because it ships with its colors, and `classic` is the theme of the
journal sets.)

A selection of twelve, on the same growth curves:

```{r gallery, fig.width = 8, fig.height = 9}
gallery_sets <- c("bw", "classic", "minimal", "economist", "wsj", "tufte",
                  "solarized", "ipsum", "nyt", "midnight", "royal", "deepblue")

gallery_base <- ggplot(growth, aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 1.2) +
  labs(x = NULL, y = NULL, color = NULL)

tiles <- lapply(gallery_sets, function(s) {
  apply_theme_set(gallery_base, s) + ggtitle(s) + theme(legend.position = "none")
})
cowplot::plot_grid(plotlist = tiles, ncol = 3)
```

A few practical notes:

- `ipsum` and `ipsum_rc` ask for the fonts "Arial Narrow" and "Roboto
  Condensed". If a font is not installed R falls back to its default family and
  the plot still draws.
- 44 of the 49 sets render markdown in their titles, legends and facet labels.
  The exceptions are `nyt`, `midnight`, `royal`, `deepblue` and `solid`. See
  [Markdown-Enabled Themes](markdown-themes.html).
- The dark sets, `midnight`, `royal`, `deepblue` and `flexoki_dark`, also set
  the default color of points, lines and text to a light one, so that a layer
  without a color mapping, such as `geom_point()`, shows on the dark panel. This
  needs ggplot2 4.0 or later; with an older version the dark sets draw those
  layers black.

---

## 4. Adjusting a Set

`apply_theme_set()` returns a normal ggplot object, and anything you add
afterwards takes precedence. Adding a second scale for the same aesthetic
replaces the one from the set (ggplot2 prints a message about it), and
`theme()` changes single elements of the set's theme:

```{r adjust}
p <- ggplot(growth, aes(time, value, color = strain)) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = "Incubation time (hr)", y = "OD600", color = NULL)

apply_theme_set(p, "fivethirtyeight") +
  scale_color_set("lancet") +
  theme(legend.position = "top")
```

This keeps the look of the `fivethirtyeight` set but swaps its three-color
palette, too small for six strains, for `lancet`, which has nine colors.

---

## 5. Sources and Credits

`themeset` does not contain the themes and palettes that it offers: it calls
the packages below by name, and the work is theirs. Please credit them when you
use their sets.

| Package | What `themeset` reaches in it | Authors |
|---------|-------------------------------|---------|
| ggplot2 | the themes `bw`, `classic`, `dark`, `light`, `linedraw`, `minimal` | the ggplot2 authors |
| ggsci | the palettes of the journal sets, and 16 more | Nan Xiao and contributors |
| ggthemes | most theme-only sets, and `fivethirtyeight` | Jeffrey B. Arnold and contributors |
| hrbrthemes | `ipsum`, `ipsum_rc` | Bob Rudis and contributors |
| tvthemes | `simpsons`, `avatar` | Ryo Nakagawara |
| cowplot | `cowplot`, `minimal_grid` | Claus O. Wilke |
| OlinkAnalyze | `olink` | Olink and the authors of OlinkAnalyze |
| flexoki | `flexoki_light`, `flexoki_dark` | Christopher T. Kenny; the Flexoki color scheme is by Steph Ango |
| ggtext | markdown in titles, labels and legends | Claus O. Wilke and Brenton M. Wiernik |
| wesanderson (optional) | `wesanderson::` palettes | Karthik Ram and Hadley Wickham |
| biopalette (optional) | `biopalette::` palettes | Yibin Zhou |
| ggpalettes (optional) | `ggpalettes::` palettes | Yaoxiang Li |

Two ideas come from packages that `themeset` does not use. The names of the
form `package::palette` follow paletteer by Emil Hvitfeldt, which reaches many
palette collections that way. The ggdark package by Neal Grantham showed that a
dark theme has to change the default color of the geoms too; `themeset` does
that with the theme element that ggplot2 4.0 provides for it, and shares no code
with ggdark.

---

## Summary

| To get | Use |
|--------|-----|
| A theme with matching colors | `apply_theme_set(p, "<complete set>")`, for example `"simpsons"`, `"nejm"` or `"flexoki_dark"` |
| Only a palette | `scale_color_set("<name>")` and `scale_fill_set("<name>")`, or `scale_color_*()` |
| The names of the palettes | `list_color_scales()` |
| Only a look | `apply_theme_set(p, "<theme-only set>")`, or call the theme directly |
| A set's look with another palette | Add a `scale_*()` after `apply_theme_set()` |
